### 2009-07-17 papers/GeneratingSensorModels/paper.tex
Wrote outline for two more papers, one on VHDL<->BON refinement, and the other on SensorML<->JML refinement via BON.

%
% "Generating Formal Executable Sensor Models from their Descriptions"
% for UNKNOWN09
% Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass{article} 
\usepackage{times}

\usepackage{ifpdf}
\usepackage{a4wide}
\usepackage{pdfsync}

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi

%\usepackage{todonotes}

% \usepackage{html}
% \usepackage{url}
\usepackage{xspace}
%\usepackage{doublespace}
\usepackage{tabularx}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage{stmaryrd}
% \ifpdf
% \usepackage[centredisplay]{diagrams}
% \else
% \usepackage[centredisplay,PostScript=dvips]{diagrams}
% \fi
\usepackage{float}

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Vieri del Bianco and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}

%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

%% \input{kt}

%=====================================================================

\begin{document}

\title{Generating Formal Executable Sensor Models\\ from their Descriptions}

\author{Joseph R. Kiniry, Vieri del Bianco, Dragan Stosic\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org, vieri.delbianco@gmail.com, dragan.stosic@gmail.com\\
}

\maketitle

%======================================================================
%\thispagestyle{empty}
\begin{abstract}

  In this work, rather than writing sensor simulators by hand, we
  automatically refine a SensorML specification through a series of
  stages to a formally specified, executable, functional
  model-based specification written in the Java Modeling Language.
  This resulting simulator implementation is completely validated with
  (for the most part, automatically generated) unit tests and is
  verified with the Extended Static Checker for Java, ESC/Java2.

\end{abstract}

%======================================================================
\section{Introduction}

The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
sensor system installed at the University College Dublin in Dublin,
Ireland.  \textbf{Our facility provides a capability---unique in the world in
terms of its scale and flexibility---for performing in-depth
investigation of both the specific and general issues in large-scale
sensor networking.}

This system integrates a sensor platform, a datastore, and a compute
farm.  The sensor platform is a custom-designed, but inexpensive,
sensor platform (called the \ST) paired with general-purpose
small-scale compute nodes, including everything from low-power PDAs to
powerful touchscreen-based portable computers.  There are over a dozen
sensors on the \ST itself, and new sensors can be added by adding them
to the USB bus.

A software-based \STS has been designed and implemented that is meant
to mimic the \ST hardware.  A part of this simulation is, of course,
simulating the sensors themselves.  The challenge in such simulation
is to ensure that a sensor simulator properly mimics the behavior of
the actual hardware sensor.  While one can specify the behavior of
such a sensor from a hardware data sheet and then hand-implement the
sensor simulator, this process is problematic on several fronts.  Not
only is one concerned about the accuracy of the simulation, but
writing and maintaining these simulations is labor-intensive.

A popular means by which to describe sensor data is to the use the
SensorML specification language.  Using SensorML one can describe
sensors and measurement processes, including an electronic
specification sheet for a sensor, the lineage of an observation, etc.
We would hope that this declarative specification is directly related
to its simulation---in other words, the simulation must be a
refinement of the SensorML specification.  But at the moment the is no
way one can define or perform this refinement other than programming
against a SensorML API by hand.

To overcome this challenge, we: (1) formally define a sensor
meta-model in the BON specification language; (2) define a formal
refinement relation between the (XML) meta-model of SensorML and this
BON meta-model; (3) using this refinement, we automatically generate
an informal and formal BON model from any SensorML model; (4) using
the Beetlz tool, from the formal BON model, we generate a formally
specified, executable, functional model-based specification written in
the Java Modeling Language, and; (5) we validate and verify, using
unit testing and extended static checking respectively, this JML
sensor model.

Consequently, using formal BON, we can define a \emph{sensor system}
as a composition of sensor models, and thus derive a complete formally
specified and verified, executable sensor system model.  To
demonstrate this process we describe an interesting slice of the \ST
specification and its simulator.

%=====================================================================
\section{Conclusion}

A conclusion.

%======================================================================
%% \nocite{ex1,ex2}
\bibliographystyle{latex8}
\bibliography{extra,%
              abbrev,%
              ads,%
              category,%
              complexity,%
              hypertext,%
              icsr,%
              knowledge,%
              languages,%
              linguistics,%
              meta,%
              metrics,%
              misc,%
              modeling,%
              modeltheory,%
              reuse,%
              rewriting,%
              softeng,%
              specification,%
              ssr,%
              technology,%
              theory,%
              web,%
              upcoming,%
              upcoming_conferences,%
              conferences,%
              workshops,%
              verification,%
              escjava,%
              jml,%
              nijmegen}

%======================================================================
% Fin

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% eval: 
%%% TeX-master: t
%%% End: