### 2009-08-19 papers/BON_VHDL/iarrayattr.tex
VHDL description

\begin{center}
\begin{tabular}{ c  c  c }
\multicolumn{3}{c}{\textbf{ATTRIBUTES OF ARRAY TYPES AND NATURES}}\\ 
\hline \hline
VHDL-AMS & BON & Description \\ 
\hline \hline
\begin{minipage}[c]{2.4cm} 
\smallskip \smallskip
\lstinlinenc|A'left(n)| \\
\lstinlinenc|A'right(n)|\\
\lstinlinenc|A'high(n)| \\
\lstinlinenc|A'low(n)| \\
\lstinlinenc|A'range(n)| \\
\lstinlinenc|A'reverse_range(n)| \\
\lstinlinenc|A'length(n)| \\
\lstinlinenc|A'ascending(n)| \\
\smallskip \smallskip
\end{minipage}  
&
\begin{minipage}[c]{4.6cm}
\centering 
\begin{lstlisting}[language=Bon]
deferred class IARRAY[T]
 feature -> n:INTEGER
  left:T -> n:INTEGER
  right:T-> n:INTEGER
  low:T  -> n:INTEGER
  high:T -> n:INTEGER
  range:INTEGER     
         -> n:INTEGER
  reverse_range:INTEGER 
         -> n:INTEGER
  length:INTEGER 
         -> n:INTEGER
  ascending:BOOLEAN 
         -> n:INTEGER
end
\end{lstlisting}
\smallskip \smallskip
\end{minipage}  
& 
\begin{minipage}[c]{6cm}  
\smallskip \smallskip
This class represents array nature interface 
with following features: leftmost value in index range of dimension n,\xspace 
rightmost value in index range of dimension n,\xspace 
least value in index range of dimension n,\xspace 
greatest value in index range of dimension n,\xspace
index range of dimension n,\xspace
index range of dimension n reversed in direction and bounds,\xspace
length of index range of dimension n,\xspace and
ordering.\xspace
\smallskip \smallskip
\end{minipage}\\ 
\end{tabular}
\end{center}