### 2009-08-19 papers/BON_VHDL/iquantity.tex

\begin{center}
\begin{tabular}{ c  c  c }
\multicolumn{3}{c}{\textbf{ATTRIBUTES OF QUANTITY}}\\ 
\hline \hline
VHDL-AMS & BON & Description \\ 
\hline \hline

\begin{minipage}[l]{2.4cm} 
\smallskip \smallskip 
\lstinlinenc|Q'tolerance|  \\
\lstinlinenc|Q'above(E)|   \\
\lstinlinenc|Q'delayed(t)| \\
\lstinlinenc|Q'dot|        \\
\lstinlinenc|Q'integ|      \\
\lstinlinenc|Q'ztf| \\
\lstinlinenc|(num,den,t| \\
\lstinlinenc|initial_delay)|       \\
\lstinlinenc|Q'ltf| \\
\lstinlinenc|(num,den)| \\
\lstinlinenc|Q'zoh| \\
\lstinlinenc|(t,initial_delay)|       \\
\lstinlinenc|Q'slew|       \\
\lstinlinenc|(max_rising_slope,| \\
\lstinlinenc|max_falling_slope)| \\
\smallskip \smallskip
\end{minipage}  
&
\begin{minipage}[c]{4.6cm}
\centering 
\smallskip \smallskip
\begin{lstlisting}[language=Bon]
deferred class IQUANTITY
 feature 
  tolerance:TOLERANCE
  above:BOOLEAN
   -> e:REAL
  delayed:REAL
   -> t:UNIVERSAL_TIME
  dot:FREE_QUANTITY 
  integ:FREE_QUANTITY
  ztf:FREE_QUANTITY
   -> num:SEQUENCE[REAL]
   -> den:SEQUENCE[REAL]
   -> t:UNIVERSAL_TIME
   -> initial_delay:UNIVERSAL_TIME 
  ltf:FREE_QUANTITY
   -> num:SEQUENCE[REAL]
   -> den:SEQUENCE[REAL] 
  zoh:FREE_QUANTITY
   -> t:UNIVERSAL_TIME
   -> initial_delay:UNIVERSAL_TIME  
  slew:FREE_QUANTITY
   -> max_rising_slope:REAL
   -> max_falling_slope:REAL
end
\end{lstlisting}
\smallskip \smallskip
\end{minipage}  
& 
\begin{minipage}[c]{6cm}  
\smallskip \smallskip
This calss represents quantity interface with following
attributes: \\
Tolerance group string presentation for scalar quantity.
Above attribute is true if the value of Q is grater than 
the value of e, otherwise false. \\
Delayed attribute represents value of quantity delayed by T.NOW-T.\\
Dot attribute represents quantity with the same type 
as quantity Q whose value is the \textbf{derivative} 
of Q with respect to time.\\
Integ attribute represents quantity with the same type 
as quantity Q whose value is the \textbf{integral} 
of Q with respect to time. \\
Ztf quantity represents \emph{z-domain} transfer function of Q.\\
Ltf quantity represents \emph{Laplace-domain} transfer function of Q. \\
Zoh quantity represents \emph{zero order hold} of Q.\\
Slow attribute represents quantity whose value follows that of Q, 
but whose derivative with respect to time is limited by the 
slopes.
\smallskip \smallskip
\end{minipage}\\ 
\end{tabular}
\end{center}