### 2009-08-19 papers/BON_VHDL/static.tex

\newcommand{\duc}{\emph{design units collection}\xspace}
\newcommand{\ducs}{\emph{design units collections}\xspace}
\newcommand{\du}{\emph{design unit}\xspace}
\begin{center}
\begin{tabular}{ c  c  c }
\multicolumn{3}{c}{\textbf{STATIC TYPES}}\\ 
\hline \hline
VHDL-AMS & BON & Description \\ 
\hline \hline
% first element
\begin{minipage}[c]{2.4cm}
\smallskip \smallskip
\centering
\ducs
\smallskip \smallskip
\end{minipage}
& 
\begin{minipage}[c]{4.6cm}
\smallskip \smallskip
\centering
\lstinlinen|SYSTEM_CHART|  
\smallskip \smallskip
\end{minipage}
& 
\begin{minipage}[c]{6cm}
         \smallskip \smallskip
	VHDL-AMS allows hierarchical structure modeling, 
	meaning that subsistems (\ducs), can be developed 
        independently. As an top-level description 
        (informal/formal), \lstinlinen|SYSTEM_CHART|
        connect all the \ducs together 
        to a complete design. Actually \lstinlinen|SYSTEM_CHART|
        can be understood as VHDL-AMS package as a way of grouping 
        a collection of related declarations that serve a common 
        purpose.
        \smallskip \smallskip
\end{minipage}\\

% next element

\begin{minipage}[c]{2.4cm}
\smallskip \smallskip
\centering
\duc
\smallskip \smallskip
\end{minipage}
& 
\begin{minipage}[c]{4.6cm}
\smallskip \smallskip
\centering
\lstinlinen|CLUSTER_CHART| 
\smallskip \smallskip
\end{minipage}
& 
\begin{minipage}[c]{6cm}
      \smallskip \smallskip	 
      An  subsistem which represent \duc.
      \smallskip \smallskip
\end{minipage}\\

% next element

\begin{minipage}[c]{2.4cm}
\smallskip \smallskip
\centering
\du
\smallskip \smallskip
\end{minipage}
& 
\begin{minipage}[c]{4.6cm}
\smallskip \smallskip
\centering
\lstinlinen|CLASS_CHART| 
\smallskip \smallskip
\end{minipage}
& 
\begin{minipage}[c]{6cm}
      \smallskip \smallskip	 
      \lstinlinen|CLASS_CHART|  describes \du. 
      It can be identify with the primary unit (\ent) 
      and secondary unit (\arch) structure in VHDL-AMS. 
      The \lstinlinen|CLASS_CHART|'s name corresponds 
      to the name of the entity in VHDL-AMS.
      \smallskip \smallskip
\end{minipage}\\

% next element

\begin{minipage}[c]{2.4cm}
\smallskip \smallskip
\centering
\arch
\smallskip \smallskip
\end{minipage}
& 
\begin{minipage}[c]{4.6cm}
\smallskip \smallskip
\centering
\lstinlinen|feature|
\smallskip \smallskip
\end{minipage}
&
\begin{minipage}[c]{6cm} 
      \smallskip \smallskip
      The feature section of the \lstinlinen|CLASS_CHART|, 
      where the mathematical transfer function of the component 
      is entered, holds the same information as the \arch section 
      of the VHDL-AMS.
      \smallskip \smallskip 
\end{minipage}\\

\end{tabular}
\end{center}