### 2009-09-04 papers/BON_VHDL/paper.tex
reorganized paper

%
% "Refining from the BON High-level Specification Language to the
% VHDL Hardware Description Language" for UNKNOWN09.
% Joseph R. Kiniry, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{ifpdf}
\usepackage{a4wide}
%\usepackage{pdfsync}
\usepackage{amsmath}
\usepackage{listings}

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi

%\usepackage{todonotes}

% \usepackage{html}
% \usepackage{url}
\usepackage{xspace}
%\usepackage{doublespace}
\usepackage{tabularx}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage{stmaryrd}
% \ifpdf
% \usepackage[centredisplay]{diagrams}
% \else
% \usepackage[centredisplay,PostScript=dvips]{diagrams}
% \fi
\usepackage{float}
\usepackage{subfig}

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\fi
\ifpdf
  \DeclareGraphicsRule{*}{mps}{*}{}
\fi
\def\figureautorefname{Fig.}

\lstset{
  language=Java,
  basicstyle=\footnotesize\sffamily,
  identifierstyle=\itshape,
  stringstyle=\footnotesize\ttfamily,
  commentstyle=\textup,
  columns=flexible,
  captionpos=b,
  morekeywords={},
  literate=
  {-}{{$\textendash$}}1
  {<}{{$\textless$}}1
  {<=}{{$\textless=$}}1
  {>}{{$\textgreater$}}2
  {>=}{{$\textgreater=$}}2
  {\\}{{$\backslash\backslash$}}1
  {'}{{$\textprimstress$}}1
}

\def\sectionautorefname{Section}

\def\lstinlinen{\lstinline[language=Bon,basicstyle=\normalsize\upshape]}

\def\lstinlinenv{\lstinline[language=VHDL,basicstyle=\normalsize\upshape]}

\def\lstinlinenc{\lstinline[language=VHDL,basicstyle=\normalsize\bfseries]}

\def\lstinlinenb{\lstinline[language=Bon,basicstyle=\normalsize\bfseries]}

\lstdefinelanguage{Bon} {
  morekeywords={require,ensure,feature,old,and,o,Q'slew,end,Result, Void},
  morekeywords={class_chart,class,delta,invariant,inherit,deferred,effective,NONE,make},
  morekeywords={explanation,query,command,constraint,exists},
  morecomment=[l]{--},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
 literate=,
  showstringspaces=false
}

\lstdefinelanguage{VHDL} {
  morekeywords={not,xnor,and,nand, nor,or,xor,integer,byte,real},
  morekeywords={character,enum,type,boolean,bit,time,null,std_ulogic},
  morekeywords={sll, srl, sla, sra, rol, ror},
  morekeywords={rem,abs,mod},
  morekeywords={terminal,T'reference, T'contribution,T'tolerance},
  morekeywords={reference,contribution},
  morekeywords={nature,subnature,electrical,magnetic,translational,translational_v},
  morekeywords={rotational,rotational_v,fluidic,thermal,radiant},
  morekeywords={voltage,angular_velocity,pressure,velocity},
  morekeywords={current,force,torque,volumetric,flow_rate},
  morekeywords={T'left,T'right,T'high,T'ascending,T'image(x),T'value(s)},
  morekeywords={left,right,high,ascending,image,value},
  morekeywords={T'pos(s),T'val(x),T'succ(x),T'pred(x)},
  morekeywords={pos,val,succ,pred},
  morekeywords={A'left(n),A'right(n),A'low(n),A'high(n)},
  morekeywords={left,right,low,high},
  morekeywords={A'range(n),A'reverse_range(n),A'length(n),A'ascending(n)},
  morekeywords={range,reverse_range,length,ascending},
  morekeywords={T'leftof(x),T'rightof(x),T'base},
  morekeywords={leftof,rightof,base},
  morekeywords={Q'tolerance, Q'above,Q'delayed,Q'dot,Q'integ},
  morekeywords={delayed,dot,integ},
  morekeywords={Q'slew,Q'ztf,Q'ltf,Q'zoh},
  morekeywords={N'across, N'through},
  morekeywords={across,through,zoh,ltf,slew,ztf},
  morekeywords={above,tolerance,group,quantity},
  morekeywords={entity,end,use,port,generic,is,library,all},
  morekeywords={architecture,of,through,begin,constant},
  morecomment=[l]{--},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
 literate=,
  showstringspaces=false
}

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}
\newcommand{\ent}{\emph{entity}\xspace}
\newcommand{\arch}{\emph{architecture}\xspace}
\newcommand{\archs}{\emph{architectures}\xspace}
\newcommand{\inv}{\emph{invariant}\xspace}
\newcommand{\undef}{\emph{undefined}\xspace}
\newcommand{\arcoss}{\emph{across}\xspace}
\newcommand{\through}{\emph{through}\xspace}
%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

%% \input{kt}

%=====================================================================

\begin{document}

\title{Refining from the BON High-level Specification Language to the
  VHDL-AMS Hardware Description Language}

\author{Joseph R. Kiniry and Dragan Stosic\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org and dragan.stosic@gmail.com\\
}

\maketitle

%======================================================================
%\thispagestyle{empty}
\begin{abstract}

The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
sensor system installed at the University College Dublin in Dublin,
Ireland.  \textbf{Our facility provides a capability---unique in the world in
terms of its scale and flexibility---for performing in-depth
investigation of both the specific and general issues in large-scale
sensor networking.}

This system integrates a sensor platform, a datastore, and a compute
farm.  The sensor platform is a custom-designed, but inexpensive,
sensor platform (called the \ST) paired with general-purpose
small-scale compute nodes, including everything from low-power PDAs to
powerful touchscreen-based portable computers.  The datastore is a
multi-terabyte scale scientific datastore into which sensor data
flows, and in which online and offline scientific computation of
sensor and other scientific data takes place.  The compute farm is a
large-scale Linux, Solaris, and OS~X-based compute farm used by
scientists for these online and offline scientific computations.

The \STs, as an embedded system, have several unusual built-in sensors
and emitters, sensor nodes can be extended via a general-purpose USB bus,
and \STs contain an FPGA that is dynamically updated on a per-experiment
basis.  \STs are custom-designed in VHDL-AMS.  Like many hardware-based systems,
 to write new software for or against the \ST architecture one either uses
the actual hardware or one runs a \STS.  But this simulator emulates a
piece of hardware, one would like to know if the simulator
\emph{actually} simulates the hardware.  We have developed an approach
through which we refine the VHDL-AMS specification up to the formal BON
specification language, and from this specification we generate an
executable, functional model-based specification in the JML language.
We have used this specification to implement a fully validated and
verified \ST simulator.

\end{abstract}

%======================================================================
\section{Introduction}
\label{sec:introduction}
The \ST sensor platform, represents heterogeneous collection of digital, analog
and mixed signal hardware circuits. Embedded system complexity
requires the need to strive for formal descriptions each component
separately in order to produce specification of the \STS.
Our aim is to be able to describe the structures of circuits,
the behavior of circuits and the formal link between structures and behaviors.
In other words, the behavioral specification of the \STS may be
represented as a \inv on the signals at the circuit's ports and may be obtained
by taking the conjunction of the \emph{invariants} that describe
behaviors of the individual components separately.
For this purpose we need  sophisticated tool and methodology
in order to be able to achieve this goal. 
The authors will illustrate in this article how BON can represent a high
level analogue mapping and will further illustrate a procedure for
transformation from VHDL-AMS to BON.\\
This article provides two main contributions.
\begin{enumerate}
\item
 It provides a solution how to formally emulate
 piece of hardware using BON-JML approach; up to date BON and JML
 have been used mostly in the pure object oriented software construction.   
\item
It relates hardware description code and refinement.
This means that if there  is a code that is "correct by construction",
then it will be derivable from BON (informal/formal) specification to 
JML implementation.
\end{enumerate}
The rest of this article is organized as follows.
In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
\autoref{sec:choiceobon} defines  BON principles and explains the
reasons why we have decided to use BON method.
In \autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON.
A comprehensive list of the mapping between BON and VHDL-AMS can be
found in the appendix. 
The central part of the our article is \autoref{sec:refVB}, in which
we will give practical aspect of refinement , applied on 
first order passive RC Low-Pass filter, in order to produce formal 
specification and software simulation.\\ 
%=====================================================================

\section{VHDL-AMS}
\label{sec:vhdlams}
IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal
high-level description language for electronic and multi-domain
systems. VHDL-AMS, as an superset of VHDL, supports modeling and
simulation of analog and mixed-signal systems. The main contribution
of VHDL-AMS is ability to describe the behaviour of complex system
trought differential and algebraic equations (DAEs). For this purpose,
VHDL-AMS introduces the following capabilities and therms:

\begin{itemize}
\item
\emph{Terminals} represents phisical connection points or circuit nodes
of a network connectivity. They are declared to be of various
\emph{natures}, which represent different energy domains of a system.
Terminals hold no values; their only role is to facilitate the formation
of conservative equation sets that, in turn, constrain the associated
\emph{branch quatities}.
\item
The \emph{quantities} denote a waveform or a time series of values.
The behaviour of the \emph{quantities} is given by the \emph{simultaneous
 statements}. They take values as a result of solving the set of simultaneous
ordinary differential and algebraic equations (DAEs). To simplify,
quantities can be understood as analog unknowns of (DAEs) that specify
analog behaviour. There are three kinds of quantity in VHDL-AMS: free,
branch and source quantities. A \emph{free quantity} can be used in
signal-flow modeling. A \emph{branch quantity} is used to model conservative energy
systems. A \emph{source quantity} is used for frequency and noise modeling.
\item
The \emph{simultaneous statements} creates one or more characteristic
equations that are solved by the dedicated simulation kernel a.k.a.
"analog solver", to model continuously valued behaviour. 
\item
The \emph{procedural statements} or \emph{procedural} express analog behaviour
using sequential statements. Actually procedural describes analog behaviour
by computing new values for quantities from the current value of quantities.
\item
The \emph{break statement} is an construct which allows to
initialize the DAE system at the beginning of a transient
simulation and to mark discontinuities during such a transient
simulation. For any concurrent break statement there is an associated process.
A break statement allows restart of the analog kernel,
than overrides quantity values and executes concurrently with the analogue model.

\end{itemize}

\section{Choice of BON as the modeling tool}
\label{sec:choiceobon}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation
for object-oriented systems  which emphasizes \emph{seamlessness},\xspace
\emph{reversibility} and software \emph{contracting}.
Let us define these principles precisely and point out why the principles
are important in VHDL-AMS simulation.\\
\textbf{Seamlessness and  Reversibility} \xspace \xspace
Seamlessness is the use of a continuous process
throughout the software lifecycle. Reversibility is the support for
both forward and backward development process: from analysis to design
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various
models consistent with the code. Keeping in sync various models, documetation
and specification from forward progress to backward and redesign, preclude
that design and implementation diverge to an inconsistent state.\\
\emph{Seamlessness} also plays an important role in relating behaviours at
different levels of abstraction. For instance, one may wish to analyze
a "lowpass-filter" at the level of individual components (a resistor and 
capacitor) and their behaviours, or at the more abstract level through
general behaviour (\inv) and using existential quantification (\texttt{exists})
to encapsulate all internal structure. Therefore capturing similarities between
the differing levels of abstractions far outweigh their inevitable differences,
making it possible to obtain a continuous process that ensures a direct mapping
between a formal specification and its software solution.\\
\textbf{Contracting} \xspace \xspace Contracting is the definition, for each
software element, of the associated semantic properties. BON use a full
assertion mechanism, allowing analysts to specify not only the structure of
a system but also its semantics ( constraints and invariants of the expected results ).
The common way to specify a program statement is to give a \emph{precondition} and
\emph{postcondition} on the state that the implementation must satisfy, assuming that
only certain program variables are changed.
In VHDL-AMS any predefined (implicite) attribute 
has defined \emph{pre} (\emph{post}) \emph{condition} trought input (output)
parameters and theirs constraints. Furthermore, any subtype (subnature)
represents restricted set of values from a base type (nature).
The condition that determines which values are in the subtype (subnature) 
may be contracted by specifying characteristic predicate which represents \inv.\\
Let us next look how \emph{simultaneous statements} are handled in contract design.
The \emph{simultaneous statement} represents analogue constraint of
continuous-time waveform. We will assume that the constraint of waveform
represents \inv. This premise deserve an informal explanation.\\
Briefly stated, analog kernel calculate quantities such
that the constraints specified by the simultaneous statements are satisfied.
More accurately, \emph{simultaneous statements} can be thought of as
algebraic equations. It means that the number of simultaneous statements
should be equal to the number of unknowns, i.e. \emph{free} and \emph{branch}
\emph{quantities}. For instance, when we reasoning about conservation of energy law,
the \emph{simultaneous statements} constrain the values of \emph{branch quantities}
in a model.
In other words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0})
around any closed path must be equal to zero and the sum of \through
quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node
(terminal) must be equal to zero. To be more precise, the results of summations 
above remains zero when a particular transformation is applied to quantities.
In general, they are \inv. \\
However, if our implementation does not satisfy conservation of energy law, 
 then the invariant will be broken. 
Another conclusion may be deducted from the example given above.\\
Intuitively \inv may be useful for testing the function of the hardware components.
In practice, one can runs simulator against the component in order to check 
the properties. Testing can then be done by checking whether the \inv hold or not.
For example, the most common problem of the real components is the glich - an electrical pulse 
of short duration that is usually the result of a design error.
The number of glitches that occur can be known by the fact how many times per day, 
the \inv was broken and reestablished. 
It is important because we cannot patch a fabricated component once we started
massive production. \\
Another advantage of BON is an easily-readable, English-based textual notation.
The current IEEE standard VHDL-AMS language reference manual tries to define
VHDL-AMS as well as possible in a descriptive way, explaining the
semantics in English.\\
Therefore, on the basis of the above, BON is the obvious choice
for the \STS.
Throughput this article we will make this more clearly. 
\section{Mapping between VHDL-AMS and BON}
\label{sec:mappingVB}

\section{Refinement from VHDL-AMS to BON }
\label{sec:refVB}
The main theme of this article is the translation from VHDL-AMS to BON using 
refinement approach.
We will refine the VHDL-AMS code up to the formal BON, and from this 
specification we will generate an executable specification in the JML language.
For this purpose, we will demonstrate our approach in BON-JML fashion trought 
low-pass filter example.
\input{lowpassRC}
We can implement the last equation (see \autoref{fig:lowpassfig}) using the \texttt{'dot} 
implicite attribute of quantity. For this purpose we introduce relation 
\texttt{$s \rightarrow d/dt \rightarrow 'dot$} . The transformation looks as follows: \\
\begin{equation}
v_{in} = v_{out} + \tau_psv_{out} 
   \begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle s \rightarrow d/dt}  \\=                                                                                                                  
   \end{array} 
v_{out} + \tau_p\cfrac{dv_{out}}{dt}
\begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle d/dt \rightarrow } {\scriptscriptstyle 'dot} \\=                                                                                                                  
   \end{array}
  v_{in} = v_{out} + \tau_pv_{out}{'dot} 
\end{equation} 
\input{lowpass}
We have not yet said what specification and refinement are.
Let us introduce these therms first.\\ 
A specification of a program component is a contract that gives some constraints 
on how the component is to behave but leaves freedom for the implementator to 
decide how the actual behavior of the component is to be realized.
Specification involves \emph{choice} between a number of alternatives; this 
relationship is one-to-many. As we have shown in \autoref{tab:rescap}, corresponding 
informal specifications represents high level of abstraction derived from VHDL-AMS code.

A refinement represents the concept that can be understud as series of 
correctness-preserving transformations from general perspective to 
a precise level of detail. Accurately, if we compare two specifications,
\texttt{S} and \texttt{S'}, then we can conclude that the \texttt{S} 
is more abstract than \texttt{S'} if any condition that we can establish 
with the first specification can also be established by second. 
It means that "\texttt{S is \emph{refined} by S'}", 
and we denote this  as \texttt{S $\rightsquigarrow$ S'}.
Let say that the first informal BON specification is denoted by 
\texttt{S} (see \autoref{tab:rescap}), and second, which is given 
in Table2  is denoted by \texttt{S'}.
In this case, the refinement relation represents an \emph{isomorphism} 
and maps the following:
%=====================================================================
\section{Conclusion}

A conclusion.

\section {Acknowledgments}
The research upon which this article is based was caried out with
sponsorship from Science Foundation Ireland via the
\textbf{UCD CASL SenseTile System} grant.

%% \nocite{ex1,ex2}
\bibliographystyle{plain}
\bibliography{extra,%
              abbrev,%
              ads,%
              category,%
              complexity,%
              hypertext,%
              icsr,%
              knowledge,%
              languages,%
              linguistics,%
              meta,%
              metrics,%
              misc,%
              modeling,%
              modeltheory,%
              reuse,%
              rewriting,%
              softeng,%
              specification,%
              ssr,%
              technology,%
              theory,%
              web,%
              upcoming,%
              upcoming_conferences,%
              conferences,%
              workshops,%
              verification,%
              escjava,%
              jml,%
              nijmegen}

%======================================================================
% Fin

%======================================================================
%\appendix
%\newpage
%\section {APPENDIX}
%\input{static}
%\input{arith}
%\input{shift}
%\input{relation}
%\input{logic}
%\input{types}
%\input{assignment}
%\input{iscaltypeattr}
%\input{tolerance}
%\input{effort}
%\input{flow}
%\input{nature}
%\input{iscalnature}
%\input{iarrayattr}
%\input{scalnat}
%\input{iterminal}
%\input{terminal}
%\input{iquantity}
%\input{quantity}
%\input{quantityport}
%\input{terminalport}
%\input{signal}
%\input{signalattr}
%\input{wait}

%======================================================================

\end{document}

%%% Local Variables:
%%% mode: latex
%%% eval:
%%% TeX-master: t
%%% End: