### 2009-09-17 papers/BON_VHDL/paper.tex
first draft

%
% "Refining from the BON High-level Specification Language to the
% VHDL Hardware Description Language" for UNKNOWN09.
% Joseph R. Kiniry, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{ifpdf}
\usepackage{a4wide}
%\usepackage{pdfsync}
\usepackage{amsmath}
\usepackage{listings}

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi

%\usepackage{todonotes}

% \usepackage{html}
% \usepackage{url}
\usepackage{xspace}
%\usepackage{doublespace}
\usepackage{tabularx}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage{stmaryrd}
% \ifpdf
% \usepackage[centredisplay]{diagrams}
% \else
% \usepackage[centredisplay,PostScript=dvips]{diagrams}
% \fi
\usepackage{float}
\usepackage{subfig}

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\fi
\ifpdf
  \DeclareGraphicsRule{*}{mps}{*}{}
\fi
\def\figureautorefname{Fig.}

\lstset{
  language=Java,
  basicstyle=\footnotesize\sffamily,
  identifierstyle=\itshape,
  stringstyle=\footnotesize\ttfamily,
  commentstyle=\textup,
  columns=flexible,
  captionpos=b,
  morekeywords={},
  literate=
  {-}{{$\textendash$}}1
  {<}{{$\textless$}}1
  {<=}{{$\textless=$}}1
  {>}{{$\textgreater$}}2
  {>=}{{$\textgreater=$}}2
  {'}{{$\textprimstress$}}1
}

\def\sectionautorefname{Section}

\def\lstinlinen{\lstinline[language=Bon,basicstyle=\normalsize\upshape]}

\def\lstinlinenv{\lstinline[language=VHDL,basicstyle=\normalsize\upshape]}

\def\lstinlinenc{\lstinline[language=VHDL,basicstyle=\normalsize\bfseries]}

\def\lstinlinenb{\lstinline[language=Bon,basicstyle=\normalsize\bfseries]}

\lstdefinelanguage{Bon} {
  morekeywords={require,ensure,feature,old,and,o,Q'slew,end,Result, Void},
  morekeywords={class_chart,class,delta,invariant,inherit,deferred,effective,NONE,make},
  morekeywords={explanation,query,command,constraint,exists,for_all,such_that,it_holds},
  morecomment=[l]{--},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
 literate=,
  showstringspaces=false
}

\lstdefinelanguage{VHDL} {
  morekeywords={not,xnor,and,nand, nor,or,xor,integer,byte,real},
  morekeywords={character,enum,type,boolean,bit,time,null,std_ulogic},
  morekeywords={sll, srl, sla, sra, rol, ror},
  morekeywords={rem,abs,mod,to},
  morekeywords={terminal,T'reference, T'contribution,T'tolerance},
  morekeywords={reference,contribution},
  morekeywords={nature,subnature,electrical,magnetic,translational,translational_v},
  morekeywords={rotational,rotational_v,fluidic,thermal,radiant},
  morekeywords={voltage,angular_velocity,pressure,velocity},
  morekeywords={current,force,torque,volumetric,flow_rate},
  morekeywords={T'left,T'right,T'high,T'ascending,T'image(x),T'value(s)},
  morekeywords={left,right,high,ascending,image,value},
  morekeywords={T'pos(s),T'val(x),T'succ(x),T'pred(x)},
  morekeywords={pos,val,succ,pred},
  morekeywords={A'left(n),A'right(n),A'low(n),A'high(n)},
  morekeywords={left,right,low,high},
  morekeywords={A'range(n),A'reverse_range(n),A'length(n),A'ascending(n)},
  morekeywords={range,reverse_range,length,ascending},
  morekeywords={T'leftof(x),T'rightof(x),T'base},
  morekeywords={leftof,rightof,base},
  morekeywords={Q'tolerance, Q'above,Q'delayed,Q'dot,Q'integ},
  morekeywords={delayed,dot,integ},
  morekeywords={Q'slew,Q'ztf,Q'ltf,Q'zoh},
  morekeywords={N'across, N'through},
  morekeywords={across,through,zoh,ltf,slew,ztf},
  morekeywords={above,tolerance,group,quantity},
  morekeywords={entity,end,use,port,generic,is,library,all},
  morekeywords={architecture,of,through,begin,constant},
  morecomment=[l]{--},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
 literate=,
  showstringspaces=false
}

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}
\newcommand{\ent}{entity\xspace}
\newcommand{\arch}{architecture\xspace}
\newcommand{\archs}{architectures\xspace}
\newcommand{\inv}{\emph{invariant}\xspace}
\newcommand{\undef}{\emph{undefined}\xspace}
\newcommand{\arcoss}{\emph{across}\xspace}
\newcommand{\through}{\emph{through}\xspace}

%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

%% \input{kt}

%=====================================================================

\begin{document}

\title{Refining from the VHDL-AMS Hardware Description Language 
to the BON High-level Specification Language}

\author{Joseph R. Kiniry and Dragan Stosic\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org and dragan.stosic@gmail.com\\
}

\maketitle

%======================================================================
%\thispagestyle{empty}
\begin{abstract}

The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose
sensor system installed at the University College Dublin in Dublin,
Ireland.  \textbf{Our facility provides a capability---unique in the world in
terms of its scale and flexibility---for performing in-depth
investigation of both the specific and general issues in large-scale
sensor networking.}

This system integrates a sensor platform, a datastore, and a compute
farm.  The sensor platform is a custom-designed, but inexpensive,
sensor platform (called the \ST) paired with general-purpose
small-scale compute nodes, including everything from low-power PDAs to
powerful touchscreen-based portable computers.  The datastore is a
multi-terabyte scale scientific datastore into which sensor data
flows, and in which online and offline scientific computation of
sensor and other scientific data takes place.  The compute farm is a
large-scale Linux, Solaris, and OS~X-based compute farm used by
scientists for these online and offline scientific computations.

The \STs, as an embedded system, have several unusual built-in sensors
and emitters, sensor nodes can be extended via a general-purpose USB bus,
and \STs contain an FPGA that is dynamically updated on a per-experiment
basis.  \STs are custom-designed in VHDL-AMS.  Like many hardware-based systems,
 to write new software for or against the \ST architecture one either uses
the actual hardware or one runs a \STS.  But this simulator emulates a
piece of hardware, one would like to know if the simulator
\emph{actually} simulates the hardware.  We have developed an approach
through which we refine the VHDL-AMS specification up to the formal BON
specification language, and from this specification we generate an
executable, functional model-based specification in the JML language.
We have used this specification to implement a fully validated and
verified \ST simulator.

\end{abstract}

%======================================================================
\section{Introduction}
\label{sec:introduction}
The \ST sensor platform, represents heterogeneous 
collection of digital, analog and mixed signal hardware circuits. 
The complexity of our platform requires the need to strive for formal 
descriptions each component separately and  behavior of their 
complex interdependencies, in order to produce executable specification.
To make sure that the sensor platform works properly we need 
to translate the VHDL-AMS code into the specification, and to refine and 
verify this specification.
In other words, our aim is to establish formal link between behavior of a given 
sensor platform, and derived specification of components  in order to implement 
fully validated and verified \ST simulator.
For this purpose, we need notation and methodology 
which will be able to satisfy the requirements given below.\\ 
First, both of them, should be in sync with  VHDL-AMS language and reference 
manual which tries to define VHDL-AMS as well as possible in a descriptive way,
explaining the semantics in English.\\
Second, specification derived from the notation should  be easy to attach to 
the rest of the system architecture\xspace(driver, microprocessor emulator, 
test-benches e.t.c.).\\
Third, methodology must aid efforts on creation of target system 
correctness criteria, such that are independent of implementation.
It means that verification of simulator model should take place on specification 
using reasoning techniques provided for algebraic formalism before coding starts.\\
Fourth, a precise semantics should  provides the basis for
automated forward-backward engineering.\\
Managing these requirements, and insuring that the platform definition 
(constraints) are preserved and verified through all levels of 
abstraction is where one becomes aware of gaps in already 
used design methodologies.
In our opinion, the BON-JML approach is in line with requirements given above.
The authors will illustrate in this article how BON can represent a high
level analogue mapping notation and will further illustrate a refinement 
technique from VHDL-AMS to BON and JML.\\
This article provides three main contributions.
\begin{enumerate}
\item
 It provides a solution how to formally emulate
 piece of hardware using BON-JML approach; up to date BON and JML
 have been used mostly in the pure object oriented software construction.   
\item
It relates hardware description code and refinement technique.
It means that if a satisfactory VHDL-AMS code exists, then it can be simulated  from the 
BON (informal/formal) specification, and JML implementation by an 
appropriate sequence of refinements.
\item
It introduces \inv as a general concept of modeling hardware 
components behavior and their interactions.
To be more precisely, the behavioral specification of the \STS may 
be represented as a \inv on the signals at the circuit's ports and 
may be obtained by taking the conjunction of the \emph{invariants} 
that describe behaviors of the individual components separately.\\ 
\end{enumerate}
The rest of this article is organized as follows.
In \autoref{sec:vhdlams} we have  presented an overview of VHDL-AMS language.
\autoref{sec:choiceobon} defines  BON principles and explains the
reasons why we have decided to use BON method.
In \autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON. 
The central part of the our article is \autoref{sec:refVB}, in which
we will give practical aspect of refinement, applied on 
first order passive RC low-pass filter, in order to produce formal 
specification and software simulation. In this section we 
explore this approach in detail.
In \autoref{sec:conclusion} the conclusions are given,
and directions of future studies are outlined. Finally, a comprehensive list 
of the mapping between BON and VHDL-AMS can be found in the appendix. \\ 
%=====================================================================

\section{VHDL-AMS}
\label{sec:vhdlams}
IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal
high-level description language for electronic and multi-domain
systems. VHDL-AMS, as an superset of VHDL, supports modeling and
simulation of analog and mixed-signal systems. The main contribution
of VHDL-AMS is ability to describe the behavior of complex system
trought differential and algebraic equations (DAEs). For this purpose,
VHDL-AMS introduces the following capabilities and therms:

\begin{itemize}
\item
\emph{Terminals} represents phisical connection points or circuit nodes
of a network connectivity. They are declared to be of various
\emph{natures}, which represent different energy domains of a system.
Terminals hold no values; their only role is to facilitate the formation
of conservative equation sets that, in turn, constrain the associated
\emph{branch qualities}.
\item
The \emph{quantities} denote a waveform or a time series of values.
The behavior of the \emph{quantities} is given by the \emph{simultaneous
 statements}. They take values as a result of solving the set of simultaneous
ordinary differential and algebraic equations (DAEs). To simplify,
quantities can be understood as analog unknowns of (DAEs) that specify
analog behaviour. There are three kinds of quantity in VHDL-AMS: free,
branch and source quantities. A \emph{free quantity} can be used in
signal-flow modeling. A \emph{branch quantity} is used to model conservative energy
systems. A \emph{source quantity} is used for frequency and noise modeling.
\item
The \emph{simultaneous statements} creates one or more characteristic
equations that are solved by the dedicated simulation kernel a.k.a.
"analog solver", to model continuously valued behavior. 
\item
The \emph{procedural statements} or \emph{procedural} express analog behaviour
using sequential statements. Actually procedural describes analog behaviour
by computing new values for quantities from the current value of quantities.
\item
The \emph{break statement} is an construct which allows to
initialize the DAE system at the beginning of a transient
simulation and to mark discontinuities during such a transient
simulation. For any concurrent break statement there is an associated process.
A break statement allows restart of the analog kernel,
than overrides quantity values and executes concurrently with the analogue model.

\end{itemize}

\section{Choice of BON as the modeling tool}
\label{sec:choiceobon}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation
for object-oriented systems  which brings \emph{seamlessness},\xspace
\emph{reversibility} and software \emph{contracting}.
Let us define these principles precisely and point out why the principles
are important in VHDL-AMS simulation.\\
\textbf{Seamlessness and  Reversibility} \xspace \xspace
Seamlessness is the use of a continuous process
throughout the software life-cycle. Reversibility is the support for
both forward and backward development process: from analysis to design
and implementation, and back.
In VHDL-AMS, modelling represents an continuous top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various
models consistent with the code. Keeping in sync various models, documentation
and specification from forward progress to backward and redesign, preclude
that design and implementation diverge to an inconsistent state.\\
\emph{Seamlessness} also plays an important role in relating behaviors at
different levels of abstraction. For instance, one may wish to analyze
a "lowpass-filter" at the level of individual components (a resistor and 
capacitor) and their behaviours, or at the more abstract level through
general behaviour (\inv).
 Therefore, capturing similarities between the differing levels of abstractions,
making it possible to obtain a continuous process that ensures a direct mapping
between a formal specification and its software solution.\\
\textbf{Contracting} \xspace \xspace Contract specification describe 
system components interaction, based on mutual obligations and benefits, 
and, for each software element, associated semantic properties. BON use a full
assertion mechanism, allowing analysts to specify both:\xspace (1) the \emph{structure} of
a system and (2)\xspace \emph{semantics} ( constraints and invariants).
The common way to specify a program statement is to give a \emph{precondition} and
\emph{postcondition} on the state that the implementation must satisfy, assuming that
only certain program variables are changed.
In VHDL-AMS any predefined (implicate) attribute 
has defined \emph{pre} (\emph{post}) \emph{condition} trought input (output)
parameters and theirs constraints. Furthermore, any subtype (subnature)
represents restricted set of values from a base type (nature).
The condition that determines which values are in the subtype (subnature) 
may be contracted by specifying characteristic predicate which represents \inv.\\
Let us next look how \emph{simultaneous statements} are handled in contract design.
The \emph{simultaneous statement} represents analogue constraint of
continuous-time waveform. We will assume that the constraint of waveform
represents \inv. This premise deserve an informal explanation.\\
Briefly stated, analog kernel calculate quantities such
that the constraints specified by the simultaneous statements are satisfied.
More accurately, \emph{simultaneous statements} can be thought of as
algebraic equations. It means that the number of simultaneous statements
should be equal to the number of unknowns (\emph{quantities}). 
For instance, when we reasoning about the conservation of energy law,
the \emph{simultaneous statements} constrain the values of \emph{branch quantities}
in a model.
In other words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across quatities =0})
around any closed path must be equal to zero and the sum of \through
quantities (\texttt{$\sum_{node}$ through quatities =0}) at any node
(terminal) must be equal to zero. To be more precise, the results of summations 
above remains zero when a particular transformation is applied to quantities.
In general, they are \inv. \\
However, if our implementation does not satisfy the conservation of energy law, 
 then the invariant will be broken. 
Another conclusion may be deducted from the example given above.\\
Intuitively \inv may be useful for testing the function of the hardware components.
In practice, one can runs simulator against the component in order to check 
the properties. Testing can then be done by checking whether the \inv hold or not.
For example, the most common problem of the real components is the glich - an electrical pulse 
of short duration that is usually the result of a design error.
The number of glitches that occur can be known by the fact how many times per day, 
the \inv was broken and reestablished. According to that we can conclude whether the 
quality of a component is satisfactory. Unlike in the case of software, where 
correction of an error is relatively inexpensive, an detected hardware error 
may require product replacement, which is associated with great expenses. 
In other words, we cannot patch a fabricated component once we started massive production. \\
Therefore, on the basis of the above, BON is the obvious choice
for the \STS.
Throughput this article we will make this more clearly. 
\section{Mapping between VHDL-AMS and BON}
\label{sec:mappingVB}
In this section, we briefly describe the mapping between VHDL-AMS and BON notation 
which is discussed in the points below.\\
\textbf{Structure} \xspace \xspace VHDL-AMS allows hierarchical structure 
modeling, meaning that the model subsystems can be developed independently.
As an top-level description, \lstinlinen|SYSTEM_CHART| in the BON notation  
can connect all the model subsystems  together to a model design.
In BON, model subsystem is denoted as the \lstinlinen|CLUSTER_CHART| whereas 
module is represented as the \lstinlinen|CLASS_CHART|.
Each module has a \texttt{port} interface: a set of \emph{inputs} and \emph{outputs} used for 
connecting modules with environment. The \texttt{port} section in 
\texttt{\ent}structure (input, output and generic values) are represented as 
a \emph{features} in the \lstinlinen|CLASS_CHART|.
The \lstinlinen|CLASS_CHART| \xspace can be equated with \ent-\arch 
structure in VHDL-AMS and the name of the \lstinlinen|CLASS_CHART|\xspace 
corresponds to the name of \texttt{\ent}in VHDL-AMS.\\ 
\textbf{Behavior} \xspace \xspace The module also contains an 
\texttt{\arch} implementation, which determines the way the module 
processes \emph{input}/\emph{output} signals and performs computations.
The \emph{constraint} section in the \lstinlinen|CLASS_CHART| 
contains the transfer function and holds the same information as 
the \texttt{\arch}section of the VHDL-AMS.\\ 
\textbf{Types and operators} \xspace \xspace There are many features of 
VHDL-AMS which are already mapped in BON by default, such as arithmetic, 
relational and logical operators and types like \xspace \texttt{integer},
\xspace \texttt{byte},\xspace\texttt{real},\xspace\texttt{character} and 
\xspace\texttt{boolean}.
Vector may be represented as \lstinlinen|SEQUENCE[TYPE]|, 
whereas record, whose elements are of heterogeneous types, can be mapped 
as \lstinlinen|SET| or \lstinlinen|TABLE|.
In BON, the \texttt{time} can be represented as the
\lstinlinen|UNIVERSAL_TIME|, in order to specify \emph{delays}. Actually 
time can be thought of as object with integer as a value and measurement 
units as a \xspace \lstinlinen|SEQUENCE| feature.     
In order to specify energy domain for nodes of circuits we have developed 
various standard \texttt{nature} templates for \texttt{terminals}.
We have introduced \xspace\lstinlinen|FREE|,\xspace \lstinlinen|BRANCH|
and  \xspace\lstinlinen|SOURCE| \space \lstinlinen|QUANTITY| objects, 
which are intended to describe signal-flow modeling, the 
conservation of energy low and frequency-noise related problems. 
There are also defined specifications for effort and 
flow domains, tolerance as well as appropriate contracts for 
scalar types(natures), array types(natures), 
terminals and quantity attributes.
Whereas the value assignment, in VHDL-AMS has the same purpose as in the BON,
the signal assignment, which schedules a new value to be applied to a 
signal later, is simulated in specification, but  denotational semantic
is not yet translated in the BON as well as for concurrency and 
the \texttt{wait} statement.

\section{Refinement from VHDL-AMS to BON }
\label{sec:refVB}
The main theme of this article is the translation from VHDL-AMS to BON using 
refinement approach.
We will refine the VHDL-AMS code up to the formal BON, and from this 
specification we will generate an executable specification in the JML language.
For this purpose, we will demonstrate our approach in BON-JML fashion trought  
low-pass filter example.\\
\subsection{Low-Pass filter transfer function }
\label{subsec:lptransfunc} 
\input{lowpassRC}
The most simple low-pass filter is passive RC network 
shown in \autoref{fig:lowpassfig}.
According to equation given above (see \autoref{fig:transfun}) 
we can produce VHDL-AMS \emph{simultaneous statement} using 
\texttt{$s = d/dt = 'dot$} relation.
The transformation looks as follows: \\
\begin{equation}
v_{in} = v_{out} + \tau_psv_{out} 
   \begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle s = d/dt}  \\=                                                                                                                  
   \end{array} 
v_{out} + \tau_p\cfrac{dv_{out}}{dt}
\begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle d/dt = } {\scriptscriptstyle 'dot} \\=                                                                                                                  
   \end{array}
  v_{in} = v_{out} + \tau_pv_{out}{'dot} 
\end{equation} 
 The last equation can be described as follows. 
The input voltage\xspace($v_{in}$)\xspace equals summation 
of output voltage\xspace($v_{out}$)\xspace and multiplication 
of the time constant\xspace($\tau_p$)\xspace of the pole and the
first derivative of the capacitor's voltage\xspace($v_{out}{'dot}$)\xspace 
over time.\\
Now we can discretize previous equation. Let us represent $v_{in}$ by a series 
of samples $v_{in(1..n)}$. At the same points in time, we can represent $v_{out}$ by 
series of sample $v_{out(1..n)}$. We can assume that samples are taken at points in 
time incrementing by $\triangle$t. Now we have equivalent equations:
\begin{equation}
v_{in(i)}- v_{out(i)}=\tau_p(\frac{v_{out(i)}- v_{out(i-1)}} {\triangle t});                                                                                                                  
\end{equation}

\begin{equation}
v_{out(i)}=v_{in(i)}(\frac{\triangle t}{\tau_p + \triangle t}) +
 v_{out(i-1)}(\frac{\tau_p}{\tau_p + \triangle t});                                                                                                                  
\end{equation}

\begin{equation}
v_{out(i)} 
 \begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle \alpha= \triangle t / (\tau_p + \triangle t)}  \\=                                                                                                           
 \end{array}
 \alpha v_{in(i)} + (1- \alpha) v_{out(i-1)};                                                                                                                  
\end{equation}

The last equation gives us a way to calculate the output samples in terms of the 
input samples and the preceding output.
But how to represent these unknowns in BON ?
We introduce the  \lstinlinen|BRANCH_QUANTITY| objects like wrappers around unknowns,
with features as fallows:
\begin{itemize}
\item
\emph{across} which represent current:$i_{out}$ 
\item 
\emph{through} which represent voltage: $v_{in}$, $v_{out}$. 
\item 
\emph{terminal sequences} which describes input/output terminal(s). 
\end{itemize}
Now we are in presence of BON notation and we may start
with the  refinement of the VHDL-AMS code.\\
\subsection{Refine example }
\label{subsec:refine}     
We have not yet said what specification and refinement are.
Let us introduce these terms first.\\
A \emph{specification} of a program component is a 
contract that gives some constraints 
on how the component is to behave but leaves freedom for the implementator to 
decide how the actual behavior of the component is to be realized. 
As we have shown in \autoref{tab:rescapfirst}, corresponding 
informal specifications represents high level of abstraction derived 
from VHDL-AMS code.\\
A \emph{refinement} represents the concept that can be understood as series of 
correctness-preserving transformations from general perspective to 
a precise level of detail. Accurately, if we compare two specifications,
\texttt{S} and \texttt{S'}, then we can conclude that the \texttt{S} 
is more abstract than \texttt{S'} if any condition that we can establish 
with the first specification can also be established by second. 
It means that "\texttt{S is \emph{refined} by S'}", 
and we denote this  as \texttt{S $\sqsubseteq$ S'}.
The \emph{refinement} which was derived in three steps, 
from informal to more formal BON and finally to JML, 
is represented using relation \texttt{$\sqsubseteq$}. This relation represent 
\emph{isomorphism} and has inverse relation \texttt{$\sqsupseteq$}. 
It means that once we have more formal specification, we can refine 
and seamlessly  translate specification in previous abstract level,  
keeping specification validity and documentation consistency. 
In other  words, we use two important BON principles to achieve 
as much as precise model consistency: the seamlessness and  
reversibility(see \autoref{sec:choiceobon}). 
Let say that the first informal BON specification is denoted by 
\texttt{S} (see \autoref{tab:rescapfirst}), and second and third,
which are given in \autoref{tab:rescapsecond}  is denoted by \texttt{S',S''}.
The forward process refinement can be represented as:\xspace
\texttt{S $\sqsubseteq$ S' $\sqsubseteq$ S''}, while the backward process
can be represented as \xspace \texttt{S''$\sqsubseteq$ S' $\sqsubseteq$ S}.  
Using all BON principles, we have developed an specification of RC low-pass
filter, and we have produced fully validated and verified low-pass filter.
\input{lowpass1}
\input{lowpass2}

%=====================================================================
\section{Conclusion}
\label{sec:conclusion}
The BON notation has originally been used for description of 
software models. However, after certain elaboration it can be used 
for description of hardware components. Practical utilization of the approach
in the project on mapping between VHDL-AMS and BON using refinement technique,
has demonstrated through low-pass filter example and shown that simulation
of component requires relatively little and inexpensive  efforts. Moreover, our article shown
that the development of the specification produced by the BON and translated 
into JML can be applied as refinement procedure: from VHDL-AMS to informal BON, 
next, from informal to more formal BON and finally, to JML specification. 
The BON support general software development and should be adapted for concrete
applications and language as VHDL-AMS. For this purpose we have developed various 
VHDL-AMS related types, operators and relations and we have defined appropriate 
specifications for all of them. However, the behavior of
\texttt{break} statement and  concurrency, for example, have not been translated 
yet in the BON notation. We have started a new phase of project named \emph{eBon} in which
the issues mentioned above, and creation of appropriate components and relations 
are main subject.

\section {Acknowledgments}
The research upon which this article is based was caried out with
sponsorship from Science Foundation Ireland via the
\textbf{UCD CASL SenseTile System} grant.

%% \nocite{ex1,ex2}
\bibliographystyle{plain}
\bibliography{extra,%
              abbrev,%
              ads,%
              category,%
              complexity,%
              hypertext,%
              icsr,%
              knowledge,%
              languages,%
              linguistics,%
              meta,%
              metrics,%
              misc,%
              modeling,%
              modeltheory,%
              reuse,%
              rewriting,%
              softeng,%
              specification,%
              ssr,%
              technology,%
              theory,%
              web,%
              upcoming,%
              upcoming_conferences,%
              conferences,%
              workshops,%
              verification,%
              escjava,%
              jml,%
              nijmegen}

%======================================================================
% Fin

%======================================================================
\appendix
\newpage
\section {APPENDIX}
\input{static}
\input{arith}
\input{shift}
\input{relation}
\input{logic}
\input{types}
\input{assignment}
\input{iscaltypeattr}
\input{tolerance}
\input{effort}
\input{flow}
\input{nature}
\input{iscalnature}
\input{iarrayattr}
\input{scalnat}
\input{iterminal}
\input{terminal}
\input{iquantity}
\input{quantity}
%\input{quantityport}
%\input{terminalport}
%\input{signal}
%\input{signalattr}
%\input{wait}
%\input{bitbon}
%======================================================================

\end{document}

%%% Local Variables:
%%% mode: latex
%%% eval:
%%% TeX-master: t
%%% End: