### 2009-11-09 papers/BON_VHDL/paper.tex
Reformatted as proper LaTeX and used fill-region to reformat syntax for easier reading and writing.  Paper has been revised and annotated by-hand by Joe.

%
% "Refining from the BON High-level Specification Language to the
% VHDL Hardware Description Language" for UNKNOWN09.
% Joseph R. Kiniry, Dragan Stosic
% $Id: paper.tex 2795 2007-09-23 19:35:53Z dmz $
%

\documentclass{article}
\usepackage{times}
\usepackage{multirow}
\usepackage{ifpdf}
\usepackage{a4wide}
%\usepackage{pdfsync}
\usepackage{amsmath}
\usepackage{listings}

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi

%\usepackage{todonotes}

% \usepackage{html}
% \usepackage{url}
\usepackage{xspace}
%\usepackage{doublespace}
\usepackage{tabularx}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage{stmaryrd}
% \ifpdf
% \usepackage[centredisplay]{diagrams}
% \else
% \usepackage[centredisplay,PostScript=dvips]{diagrams}
% \fi
\usepackage{float}
\usepackage{subfig}

\ifpdf
\usepackage[pdftex,bookmarks=false,a4paper=false,
            plainpages=false,naturalnames=true,
            colorlinks=true,pdfstartview=FitV,
            linkcolor=blue,citecolor=blue,urlcolor=blue,
            pdfauthor="Joseph R. Kiniry and Dragan Stosic"]{hyperref}
\else
\usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\fi
\ifpdf
  \DeclareGraphicsRule{*}{mps}{*}{}
\fi
\def\figureautorefname{Fig.}

\lstset{
  language=Java,
  basicstyle=\footnotesize\sffamily,
  identifierstyle=\itshape,
  stringstyle=\footnotesize\ttfamily,
  commentstyle=\textup,
  columns=flexible,
  captionpos=b,
  morekeywords={},
  literate=
  {-}{{$\textendash$}}1
  {<}{{$\textless$}}1
  {<=}{{$\textless=$}}1
  {>}{{$\textgreater$}}2
  {>=}{{$\textgreater=$}}2
  {'}{{$\textprimstress$}}1
}

\def\sectionautorefname{Section}

\def\lstinlinen{\lstinline[language=Bon,basicstyle=\normalsize\upshape]}

\def\lstinlinenv{\lstinline[language=VHDL,basicstyle=\normalsize\upshape]}

\def\lstinlinenc{\lstinline[language=VHDL,basicstyle=\normalsize\bfseries]}

\def\lstinlinenb{\lstinline[language=Bon,basicstyle=\normalsize\bfseries]}

\lstdefinelanguage{Bon} {
  morekeywords={require,ensure,feature,old,and,o,Q'slew,end,Result, Void},
  morekeywords={class_chart,class,delta,invariant,inherit,deferred,effective,NONE,make},
  morekeywords={explanation,query,command,constraint,exists,for_all,such_that,it_holds},
  morecomment=[l]{--},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
 literate=,
  showstringspaces=false
}

\lstdefinelanguage{VHDL} {
  morekeywords={not,xnor,and,nand, nor,or,xor,integer,byte,real},
  morekeywords={character,enum,type,boolean,bit,time,null,std_ulogic},
  morekeywords={sll, srl, sla, sra, rol, ror},
  morekeywords={rem,abs,mod,to},
  morekeywords={terminal,T'reference, T'contribution,T'tolerance},
  morekeywords={reference,contribution},
  morekeywords={nature,subnature,electrical,magnetic,translational,translational_v},
  morekeywords={rotational,rotational_v,fluidic,thermal,radiant},
  morekeywords={voltage,angular_velocity,pressure,velocity},
  morekeywords={current,force,torque,volumetric,flow_rate},
  morekeywords={T'left,T'right,T'high,T'ascending,T'image(x),T'value(s)},
  morekeywords={left,right,high,ascending,image,value},
  morekeywords={T'pos(s),T'val(x),T'succ(x),T'pred(x)},
  morekeywords={pos,val,succ,pred},
  morekeywords={A'left(n),A'right(n),A'low(n),A'high(n)},
  morekeywords={left,right,low,high},
  morekeywords={A'range(n),A'reverse_range(n),A'length(n),A'ascending(n)},
  morekeywords={range,reverse_range,length,ascending},
  morekeywords={T'leftof(x),T'rightof(x),T'base},
  morekeywords={leftof,rightof,base},
  morekeywords={Q'tolerance, Q'above,Q'delayed,Q'dot,Q'integ},
  morekeywords={delayed,dot,integ},
  morekeywords={Q'slew,Q'ztf,Q'ltf,Q'zoh},
  morekeywords={N'across, N'through},
  morekeywords={across,through,zoh,ltf,slew,ztf},
  morekeywords={above,tolerance,group,quantity},
  morekeywords={entity,end,use,port,generic,is,library,all},
  morekeywords={architecture,of,through,begin,constant},
  morecomment=[l]{--},
  morecomment=[s]{/*}{*/},
  morestring=[b]",
 literate=,
  showstringspaces=false
}

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}
\newcommand{\ent}{entity\xspace}
\newcommand{\arch}{architecture\xspace}
\newcommand{\archs}{architectures\xspace}
\newcommand{\inv}{\emph{invariant}\xspace}
\newcommand{\undef}{\emph{undefined}\xspace}
\newcommand{\arcoss}{\emph{across}\xspace}
\newcommand{\through}{\emph{through}\xspace}

%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

%% \input{kt}

%=====================================================================

\begin{document}

\title{Refining from the VHDL-AMS Hardware Description Language 
to the BON High-level Specification Language}

\author{Joseph R. Kiniry and Dragan Stosic\\
UCD CASL: Complex and Adaptive Systems Laboratory and\\
School of Computer Science and Informatics,\\
University College Dublin,\\
Belfield, Dublin 4, Ireland,\\
kiniry@acm.org and dragan.stosic@gmail.com\\
}

\maketitle

%======================================================================
%\thispagestyle{empty}
\begin{abstract}

The \textbf{UCD CASL SenseTile System} is a large-scale,
general-purpose sensor system installed at the University College
Dublin in Dublin, Ireland.  \textbf{Our facility provides a
capability---unique in the world in terms of its scale and
flexibility---for performing in-depth investigation of both the
specific and general issues in large-scale sensor networking.}

This system integrates a sensor platform, a datastore, and a compute
farm.  The sensor platform is a custom-designed, but inexpensive,
sensor platform (called the \ST) paired with general-purpose
small-scale compute nodes, including everything from low-power PDAs to
powerful touchscreen-based portable computers.  The datastore is a
multi-terabyte scale scientific datastore into which sensor data
flows, and in which online and offline scientific computation of
sensor and other scientific data takes place.  The compute farm is a
large-scale Linux, Solaris, and OS~X-based compute farm used by
scientists for these online and offline scientific computations.

The \STs, as an embedded system, have several unusual built-in sensors
and emitters, sensor nodes can be extended via a general-purpose USB
bus, and \STs contain an FPGA that is dynamically updated on a
per-experiment basis.  \STs are custom-designed in VHDL-AMS.  Like
many hardware-based systems, to write new software for or against the
\ST architecture one either uses the actual hardware or one runs a
\STS.  But this simulator emulates a piece of hardware, one would like
to know if the simulator \emph{actually} simulates the hardware.  We
have developed an approach through which we refine the VHDL-AMS
specification up to the formal BON specification language, and from
this specification we generate an executable, functional model-based
specification in the JML language.  We have used this specification to
implement a fully validated and verified \ST simulator.

\end{abstract}

%======================================================================
\section{Introduction}
\label{sec:introduction}

The \ST sensor platform, represents heterogeneous collection of
digital, analog and mixed signal hardware circuits.  The complexity of
our platform requires the need to strive for formal descriptions each
component separately and behavior of their complex interdependencies,
in order to produce executable specification.\footnote{This is not a
  good reason for performing this research.  Complexity is often the
  reason to use \emph{less} formality, nor more.  Also, why must the
  specification be \emph{executable}?}  To make sure that the sensor
platform works properly we need to translate the VHDL-AMS code into
the specification, and to refine and verify this
specification.\footnote{Too vague a statement---what are the
  \emph{real} reasons we are doing this work?}  In other words, our
aim is to establish formal link between behavior of a given sensor
platform, and derived specification of components in order to
implement fully validated and verified \ST simulator.  For this
purpose, we need notation and methodology which will be able to
satisfy the requirements given below.

First, both of them, should be in sync with VHDL-AMS language and
reference manual which tries to define VHDL-AMS as well as possible in
a descriptive way, explaining the semantics in English.

Second, specification derived from the notation should be easy to
attach\footnote{Meaning?} to the rest of the system
architecture\xspace(driver, microprocessor emulator, test-benches
e.t.c.).

Third, methodology must aid efforts on creation of target system
correctness criteria, such that are independent of implementation.  It
means that verification of simulator model should take place on
specification using reasoning techniques provided for algebraic
formalism before coding starts.

Fourth, a precise semantics should provides the basis for automated
forward-backward engineering.

Managing these requirements, and insuring that the platform definition
(constraints) are preserved and verified through all levels of
abstraction is where one becomes aware of gaps in already used design
methodologies.  In our opinion, the BON-JML approach is in line with
requirements given above.  The authors will illustrate in this article
how BON can represent a high level analogue mapping notation and will
further illustrate a refinement technique from VHDL-AMS to BON and
JML.

This article provides three main contributions.
\begin{enumerate}
\item It provides a solution how to formally emulate piece of hardware
using BON-JML approach; up to date BON and JML have been used mostly
in the pure object oriented software construction.
\item It relates hardware description code and refinement technique.
It means that if a satisfactory VHDL-AMS code exists, then it can be
simulated from the BON (informal/formal) specification, and JML
implementation by an appropriate sequence of refinements.
\item It introduces \inv as a general concept of modeling hardware
components behavior and their interactions.  To be more precisely, the
behavioral specification of the \STS may be represented as a \inv on
the signals at the circuit's ports and may be obtained by taking the
conjunction of the \emph{invariants} that describe behaviors of the
individual components separately.

\end{enumerate} The rest of this article is organized as follows.  In
\autoref{sec:vhdlams} we have presented an overview of VHDL-AMS
language.  \autoref{sec:choiceobon} defines BON principles and
explains the reasons why we have decided to use BON method.  In
\autoref{sec:mappingVB} we will show mapping between VHDL-AMS and BON.
The central part of the our article is \autoref{sec:refVB}, in which
we will give practical aspect of refinement, applied on first order
passive RC low-pass filter, in order to produce formal specification
and software simulation. In this section we explore this approach in
detail.  In \autoref{sec:conclusion} the conclusions are given, and
directions of future studies are outlined. Finally, a comprehensive
list of the mapping between BON and VHDL-AMS can be found in the
appendix.

%=====================================================================

\section{VHDL-AMS}
\label{sec:vhdlams} IEEE VHDL-AMS 1076.1 is the industry standard
mixed-signal high-level description language for electronic and
multi-domain systems. VHDL-AMS, as an superset of VHDL, supports
modeling and simulation of analog and mixed-signal systems. The main
contribution of VHDL-AMS is ability to describe the behavior of
complex system trought differential and algebraic equations
(DAEs). For this purpose, VHDL-AMS introduces the following
capabilities and therms:

\begin{itemize}
\item \emph{Terminals} represents phisical connection points or
circuit nodes of a network connectivity. They are declared to be of
various \emph{natures}, which represent different energy domains of a
system.  Terminals hold no values; their only role is to facilitate
the formation of conservative equation sets that, in turn, constrain
the associated \emph{branch qualities}.
\item The \emph{quantities} denote a waveform or a time series of
values.  The behavior of the \emph{quantities} is given by the
\emph{simultaneous statements}. They take values as a result of
solving the set of simultaneous ordinary differential and algebraic
equations (DAEs). To simplify, quantities can be understood as analog
unknowns of (DAEs) that specify analog behaviour. There are three
kinds of quantity in VHDL-AMS: free, branch and source quantities. A
\emph{free quantity} can be used in signal-flow modeling. A
\emph{branch quantity} is used to model conservative energy systems. A
\emph{source quantity} is used for frequency and noise modeling.
\item The \emph{simultaneous statements} creates one or more
characteristic equations that are solved by the dedicated simulation
kernel a.k.a.  "analog solver", to model continuously valued behavior.
\item The \emph{procedural statements} or \emph{procedural} express
analog behaviour using sequential statements. Actually procedural
describes analog behaviour by computing new values for quantities from
the current value of quantities.
\item The \emph{break statement} is an construct which allows to
initialize the DAE system at the beginning of a transient simulation
and to mark discontinuities during such a transient simulation. For
any concurrent break statement there is an associated process.  A
break statement allows restart of the analog kernel, than overrides
quantity values and executes concurrently with the analogue model.

\end{itemize}

\section{Choice of BON as the modeling tool}
\label{sec:choiceobon} %Brief description and motivation.  The
Business Object Notation (BON) is an analysis and design notation for
object-oriented systems which brings \emph{seamlessness},\xspace
\emph{reversibility} and software \emph{contracting}.  Let us define
these principles precisely and point out why the principles are
important in VHDL-AMS simulation.

\textbf{Seamlessness and Reversibility} \xspace \xspace Seamlessness
is the use of a continuous process throughout the software
life-cycle. Reversibility is the support for both forward and backward
development process: from analysis to design and implementation, and
back.  In VHDL-AMS, modelling represents an continuous
top-down-bottom-up process. Top-down approach serves for design while
bottom-up serves for verification. The possibility of seamlessly
translating changes made during a certain development phase back into
earlier phases, facilitates a stable and predictable environment and
ensures the various models consistent with the code. Keeping in sync
various models, documentation and specification from forward progress
to backward and redesign, preclude that design and implementation
diverge to an inconsistent state.

\emph{Seamlessness} also plays an important role in relating behaviors
at different levels of abstraction. For instance, one may wish to
analyze a "lowpass-filter" at the level of individual components (a
resistor and capacitor) and their behaviours, or at the more abstract
level through general behaviour (\inv).  Therefore, capturing
similarities between the differing levels of abstractions, making it
possible to obtain a continuous process that ensures a direct mapping
between a formal specification and its software solution.

\textbf{Contracting} \xspace \xspace Contract specification describe
system components interaction, based on mutual obligations and
benefits, and, for each software element, associated semantic
properties. BON use a full assertion mechanism, allowing analysts to
specify both:\xspace (1) the \emph{structure} of a system and
(2)\xspace \emph{semantics} ( constraints and invariants).  The common
way to specify a program statement is to give a \emph{precondition}
and \emph{postcondition} on the state that the implementation must
satisfy, assuming that only certain program variables are changed.  In
VHDL-AMS any predefined (implicate) attribute has defined \emph{pre}
(\emph{post}) \emph{condition} trought input (output) parameters and
theirs constraints. Furthermore, any subtype (subnature) represents
restricted set of values from a base type (nature).  The condition
that determines which values are in the subtype (subnature) may be
contracted by specifying characteristic predicate which represents
\inv.

Let us next look how \emph{simultaneous statements} are handled in
contract design.  The \emph{simultaneous statement} represents
analogue constraint of continuous-time waveform. We will assume that
the constraint of waveform represents \inv. This premise deserve an
informal explanation.

Briefly stated, analog kernel calculate quantities such that the
constraints specified by the simultaneous statements are satisfied.
More accurately, \emph{simultaneous statements} can be thought of as
algebraic equations. It means that the number of simultaneous
statements should be equal to the number of unknowns
(\emph{quantities}).  For instance, when we reasoning about the
conservation of energy law, the \emph{simultaneous statements}
constrain the values of \emph{branch quantities} in a model.  In other
words, the sum of \arcoss quantities (\texttt{$\sum_{loop}$ across
quatities =0}) around any closed path must be equal to zero and the
sum of \through quantities (\texttt{$\sum_{node}$ through quatities
=0}) at any node (terminal) must be equal to zero. To be more precise,
the results of summations above remains zero when a particular
transformation is applied to quantities.  In general, they are \inv.

However, if our implementation does not satisfy the conservation of
energy law, then the invariant will be broken.  Another conclusion may
be deducted from the example given above.

Intuitively \inv may be useful for testing the function of the
hardware components.  In practice, one can runs simulator against the
component in order to check the properties. Testing can then be done
by checking whether the \inv hold or not.  For example, the most
common problem of the real components is the glich - an electrical
pulse of short duration that is usually the result of a design error.
The number of glitches that occur can be known by the fact how many
times per day, the \inv was broken and reestablished. According to
that we can conclude whether the quality of a component is
satisfactory. Unlike in the case of software, where correction of an
error is relatively inexpensive, an detected hardware error may
require product replacement, which is associated with great expenses.
In other words, we cannot patch a fabricated component once we started
massive production.

Therefore, on the basis of the above, BON is the obvious choice for
the \STS.  Throughput this article we will make this more clearly.
\section{Mapping between VHDL-AMS and BON}
\label{sec:mappingVB} In this section, we briefly describe the mapping
between VHDL-AMS and BON notation which is discussed in the points
below.

\textbf{Structure} \xspace \xspace VHDL-AMS allows hierarchical
structure modeling, meaning that the model subsystems can be developed
independently.  As an top-level description, \lstinlinen|SYSTEM_CHART|
in the BON notation can connect all the model subsystems together to a
model design.  In BON, model subsystem is denoted as the
\lstinlinen|CLUSTER_CHART| whereas module is represented as the
\lstinlinen|CLASS_CHART|.  Each module has a \texttt{port} interface:
a set of \emph{inputs} and \emph{outputs} used for connecting modules
with environment. The \texttt{port} section in \texttt{\ent}structure
(input, output and generic values) are represented as a
\emph{features} in the \lstinlinen|CLASS_CHART|.  The
\lstinlinen|CLASS_CHART| \xspace can be equated with \ent-\arch
structure in VHDL-AMS and the name of the
\lstinlinen|CLASS_CHART|\xspace corresponds to the name of
\texttt{\ent}in VHDL-AMS.

\textbf{Behavior} \xspace \xspace The module also contains an
\texttt{\arch} implementation, which determines the way the module
processes \emph{input}/\emph{output} signals and performs
computations.  The \emph{constraint} section in the
\lstinlinen|CLASS_CHART| contains the transfer function and holds the
same information as the \texttt{\arch}section of the VHDL-AMS.

\textbf{Types and operators} \xspace \xspace There are many features
of VHDL-AMS which are already mapped in BON by default, such as
arithmetic, relational and logical operators and types like \xspace
\texttt{integer}, \xspace
\texttt{byte},\xspace\texttt{real},\xspace\texttt{character} and
\xspace\texttt{boolean}.  Vector may be represented as
\lstinlinen|SEQUENCE[TYPE]|, whereas record, whose elements are of
heterogeneous types, can be mapped as \lstinlinen|SET| or
\lstinlinen|TABLE|.  In BON, the \texttt{time} can be represented as
the \lstinlinen|UNIVERSAL_TIME|, in order to specify
\emph{delays}. Actually time can be thought of as object with integer
as a value and measurement units as a \xspace \lstinlinen|SEQUENCE|
feature.  In order to specify energy domain for nodes of circuits we
have developed various standard \texttt{nature} templates for
\texttt{terminals}.  We have introduced
\xspace\lstinlinen|FREE|,\xspace \lstinlinen|BRANCH| and
\xspace\lstinlinen|SOURCE| \space \lstinlinen|QUANTITY| objects, which
are intended to describe signal-flow modeling, the conservation of
energy low and frequency-noise related problems.  There are also
defined specifications for effort and flow domains, tolerance as well
as appropriate contracts for scalar types(natures), array
types(natures), terminals and quantity attributes.  Whereas the value
assignment, in VHDL-AMS has the same purpose as in the BON, the signal
assignment, which schedules a new value to be applied to a signal
later, is simulated in specification, but denotational semantic is not
yet translated in the BON as well as for concurrency and the
\texttt{wait} statement.

\section{Refinement from VHDL-AMS to BON }
\label{sec:refVB} The main theme of this article is the translation
from VHDL-AMS to BON using refinement approach.  We will refine the
VHDL-AMS code up to the formal BON, and from this specification we
will generate an executable specification in the JML language.  For
this purpose, we will demonstrate our approach in BON-JML fashion
trought low-pass filter example.

\subsection{Low-Pass filter transfer function }
\label{subsec:lptransfunc} 
\input{lowpassRC}

The most simple low-pass filter is passive RC network shown in
\autoref{fig:lowpassfig}.  According to equation given above (see
\autoref{fig:transfun}) we can produce VHDL-AMS \emph{simultaneous
  statement} using \texttt{$s = d/dt = 'dot$} relation.  The
transformation looks as follows:
\begin{equation}
v_{in} = v_{out} + \tau_psv_{out} 
   \begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle s = d/dt}  \\=                                                                                                                  
   \end{array} 
v_{out} + \tau_p\cfrac{dv_{out}}{dt}
\begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle d/dt = } {\scriptscriptstyle 'dot} \\=                                                                                                                  
   \end{array}
  v_{in} = v_{out} + \tau_pv_{out}{'dot} 
\end{equation}

The last equation can be described as follows.  The input
voltage\xspace($v_{in}$)\xspace equals summation of output
voltage\xspace($v_{out}$)\xspace and multiplication of the time
constant\xspace($\tau_p$)\xspace of the pole and the first derivative
of the capacitor's voltage\xspace($v_{out}{'dot}$)\xspace over time.

Now we can discretize previous equation. Let us represent $v_{in}$ by
a series of samples $v_{in(1..n)}$. At the same points in time, we can
represent $v_{out}$ by series of sample $v_{out(1..n)}$. We can assume
that samples are taken at points in time incrementing by
$\triangle$t.

Now we have equivalent equations:
\begin{equation}
v_{in(i)}- v_{out(i)}=\tau_p(\frac{v_{out(i)}- v_{out(i-1)}} {\triangle t});                                                                                          \end{equation}

\begin{equation}
v_{out(i)}=v_{in(i)}(\frac{\triangle t}{\tau_p + \triangle t}) +
 v_{out(i-1)}(\frac{\tau_p}{\tau_p + \triangle t});                                                                                                                  
\end{equation}

\begin{equation}
v_{out(i)} 
 \begin{array}[b]{c}                                                                                                                                    
      {\scriptscriptstyle \alpha= \triangle t / (\tau_p + \triangle t)}  \\=                                                                              \end{array}
 \alpha v_{in(i)} + (1- \alpha) v_{out(i-1)};                                                                                                                  
\end{equation}

The last equation gives us a way to calculate the output samples in
terms of the input samples and the preceding output.  But how to
represent these unknowns in BON ?  We introduce the
\lstinlinen|BRANCH_QUANTITY| objects like wrappers around unknowns,
with features as follows:
\begin{itemize}
\item
\emph{across} which represent current:$i_{out}$ 
\item 
\emph{through} which represent voltage: $v_{in}$, $v_{out}$. 
\item 
\emph{terminal sequences} which describes input/output terminal(s). 
\end{itemize}
Now we are in presence of BON notation and we may start with the
refinement of the VHDL-AMS code.

\subsection{Refine example }
\label{subsec:refine}     
We have not yet said what specification and refinement are.  Let us
introduce these terms first.

A \emph{specification} of a program component is a contract that gives
some constraints on how the component is to behave but leaves freedom
for the implementator to decide how the actual behavior of the
component is to be realized.  As we have shown in
\autoref{tab:rescapfirst}, corresponding informal specifications
represents high level of abstraction derived from VHDL-AMS code.

A \emph{refinement} represents the concept that can be understood as
series of correctness-preserving transformations from general
perspective to a precise level of detail. Accurately, if we compare
two specifications, \texttt{S} and \texttt{S'}, then we can conclude
that the \texttt{S} is more abstract than \texttt{S'} if any condition
that we can establish with the first specification can also be
established by second.  It means that "\texttt{S is \emph{refined} by
S'}", and we denote this as \texttt{S $\sqsubseteq$ S'}.  The
\emph{refinement} which was derived in three steps, from informal to
more formal BON and finally to JML, is represented using relation
\texttt{$\sqsubseteq$}. This relation represent \emph{isomorphism} and
has inverse relation \texttt{$\sqsupseteq$}.  It means that once we
have more formal specification, we can refine and seamlessly translate
specification in previous abstract level, keeping specification
validity and documentation consistency.  In other words, we use two
important BON principles to achieve as much as precise model
consistency: the seamlessness and reversibility(see
\autoref{sec:choiceobon}).  Let say that the first informal BON
specification is denoted by \texttt{S} (see
\autoref{tab:rescapfirst}), and second and third, which are given in
\autoref{tab:rescapsecond} is denoted by \texttt{S',S''}.  The forward
process refinement can be represented as:\xspace \texttt{S
$\sqsubseteq$ S' $\sqsubseteq$ S''}, while the backward process can be
represented as \xspace \texttt{S''$\sqsubseteq$ S' $\sqsubseteq$ S}.
Using all BON principles, we have developed an specification of RC
low-pass filter, and we have produced fully validated and verified
low-pass filter.  \input{lowpass1} \input{lowpass2}

%=====================================================================
\section{Conclusion}
\label{sec:conclusion} The BON notation has originally been used for
description of software models. However, after certain elaboration it
can be used for description of hardware components. Practical
utilization of the approach in the project on mapping between VHDL-AMS
and BON using refinement technique, has demonstrated through low-pass
filter example and shown that simulation of component requires
relatively little and inexpensive efforts. Moreover, our article shown
that the development of the specification produced by the BON and
translated into JML can be applied as refinement procedure: from
VHDL-AMS to informal BON, next, from informal to more formal BON and
finally, to JML specification.  The BON support general software
development and should be adapted for concrete applications and
language as VHDL-AMS. For this purpose we have developed various
VHDL-AMS related types, operators and relations and we have defined
appropriate specifications for all of them. However, the behavior of
\texttt{break} statement and concurrency, for example, have not been
translated yet in the BON notation. We have started a new phase of
project named \emph{eBon} in which the issues mentioned above, and
creation of appropriate components and relations are main subject.

\section {Acknowledgments} The research upon which this article is
based was caried out with sponsorship from Science Foundation Ireland
via the \textbf{UCD CASL SenseTile System} grant.

%% \nocite{ex1,ex2}
\bibliographystyle{plain}
\bibliography{extra,%
              abbrev,%
              ads,%
              category,%
              complexity,%
              hypertext,%
              icsr,%
              knowledge,%
              languages,%
              linguistics,%
              meta,%
              metrics,%
              misc,%
              modeling,%
              modeltheory,%
              reuse,%
              rewriting,%
              softeng,%
              specification,%
              ssr,%
              technology,%
              theory,%
              web,%
              upcoming,%
              upcoming_conferences,%
              conferences,%
              workshops,%
              verification,%
              escjava,%
              jml,%
              nijmegen}

%======================================================================
% Fin

%======================================================================
\appendix
\newpage
\section {APPENDIX}
\input{static}
\input{arith}
\input{shift}
\input{relation}
\input{logic}
\input{types}
\input{assignment}
\input{iscaltypeattr}
\input{tolerance}
\input{effort}
\input{flow}
\input{nature}
\input{iscalnature}
\input{iarrayattr}
\input{scalnat}
\input{iterminal}
\input{terminal}
\input{iquantity}
\input{quantity}
%\input{quantityport}
%\input{terminalport}
%\input{signal}
%\input{signalattr}
%\input{wait}
%\input{bitbon}
%======================================================================

\end{document}

%%% Local Variables:
%%% mode: latex
%%% eval:
%%% TeX-master: t
%%% End: