### 2010-04-11 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

\documentclass[]{final_report}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{listings}

%%%%%%%%%%%%%%%%%%%%%%
%%% Input project details
\def\studentname{Ciaran Palmer}
\def\projecttitle{SensorML on SenseTile}
\def\supervisorname{Dr. Joseph Kiniry}
\def\moderatorname{-}

\begin{document}

\maketitle
\tableofcontents\pdfbookmark[0]{Table of Contents}{toc}\newpage

%%%%%%%%%%%%%%%%%%%%%%
%%% Your Abstract here

\begin{abstract}
This thesis evaluates the use of SensorML and SOS in the development of Sensor Systems. The SenseTile system is used as a case study for this evaluation. A SenseTile web service is developed to access SenseTile sensor data based on a SensorML model of SenseTile. Formalizing senor descriptions is also looked using BON to SensorML mapping.

\end{abstract}

\newpage

%%%%%%%%%%%%%%%%%%%%%%
%%% Acknowledgments

\chapter*{Acknowledgments}

Vieri del Bianco , Dragan Stosic and Joseph R. Kiniry.

%%%%%%%%%%%%%%%%%%%%%%
%%% Introduction

\chapter{Introduction}

Sensor systems are increasingly deployed to provide sensing information for a myriad of applications. These sensor systems contain many different sensor types and are developed by communities working in different domains. The heterogeneity of these systems means that interoperability is difficult and expensive. Reliably processing sensor observations retrieved from these heterogeneous systems is also problematic. A standard way to retrieve sensor data and a standard way to describe sensors are needed to address the current problem of sensor system heterogeneity.

Standards are being developed to address the above issues. Two such standards are the Sensor Model Language (SensorML)\cite{SMLref} specification and the Sensor Observation Service (SOS)\cite{SOSref} specification. SensorML is a proposed standard for sensor description and processing of sensor observations. SensorML specifies models and XML Schema for describing sensors, the process of sensor measurement and the post processing of sensor observations. The SOS specification is a proposed standard for retrieving sensor information and sensor observations using Web Services.

The SensorML and SOS specifications are developed by the Open Geospatial Consortium (OGC) as part of the Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}. The members of the OGC have the goal of improving interoperability of sensor systems by developing open standards such as the SensorML and SOS specifications. This thesis investigates the use of these SWE standards in developing sensor systems with emphasis on the use of SensorML and SOS specifications.

Developing robust sensor systems is also a natural concern. There are now formal verification based software development processes \cite{Kiniryref} with the goal of developing very robust systems.  Currently there is no widely available formal definition of sensors or sensor data processing. Formal sensor descriptions allows sensor system development with such a verification based process and their development is examined also in this thesis.

A real sensor system called SenseTile[ref] is used as a case study in this thesis. SenseTile is a sensor and processing package that contains motion sensors, temperature sensors, audio sensors, pressure sensors and light level sensors among others. It is a replacement for standard ceiling tiles to provide smart building services as part of a Web Sensor Network.

This thesis describes a  SensorML model of SenseTile, in particular the SenseTile Sensor Board . Based on this model a SenseTile Web Service prototype is developed to allow the retrieval of the SenseTile sensor descriptions and sensor observations. The Web Service is based on the SOS specification. An assessment of SensorML and SOS specifications is provided based on this work.

Buisness Object Notatation(BON)\cite{BONref} isi used to formalize SensorML sensor descriptions in this thesis. BON is a notation and a method for Object Oriented systems analysis and design. The BON to SensorML mapping described in thesis allows SensorML sensor descriptions to then be refined to Java as described in \cite{Kiniryref} as part of developing a robust Sensor System.

\chapter{ Background Research}

\section{Sensor Web Enablement}\label{SWESec}
As described in the introduction the SensorML and SOS specifications are developed as part of the Open Geospatial Consortium (OGC) Sensor Web Enablement (SWE) initiative \footnote{http://www.opengeospatial.org/projects/groups/sensorweb}. This initiative is establishing interfaces and protocols to enable a standardized Sensor Web. It is hoped that this allows for easier integration of sensor applications across sensor networks. The OGC SWE provides a framework of open standards for use with Web-connected sensors and sensor systems. There are seven main specifications currently:

\begin{itemize}
\item  SensorModel Language(SensorML) - models and schema for sensor systems and processes surrounding measurements
\item  Sensor Observation Service (SOS) - standard web interface for accessing observations
\item  Observations \& Measurements (O\&M) - models and schema for packaging observation values
\item  Transducer Markup Language (TML) - models and schema for multiplexed data from sensor systems
\item  Sensor Planning Service (SPS) - standard web interface for tasking sensor systems and model and requesting acquisitions
\item  Sensor Alert Service (SAS) - standard web interface for publishing and subscribing to sensor alerts
\item  Web Notification Service (WNS) - standard web interface for asynchronous notification
\end{itemize}

SWE services based on the aboved specifications allow the discovery of sensors, access to the sensor data, as well as subscription to alerts, and tasking of sensors to control observations. For a good overview of SWE and it's use in Sensor Web Networks see \cite{SWEArchref}.  Along with with the SensorML and SOS specification, the O\&M specification is also used in developing the SenseTile Web Service as it defines the encoding of sensor observations.

\section{Sensor Model Language}\label{SMLsection}
Sensor Modeling Language (SensorML) is primarily used to describe sensor systems and the processing of observations from sensor systems. The SensorML specification\cite{SMLref} provides a functional model of sensors and an XML encoding to describe sensors and their observations. Some of the main uses of SensorML described in the specification are:
\begin{itemize}
\item Support sensor and sensor observation discovery.
\item Support processing of the sensor observations.
\item Support geolocation information about sensors.
\item Provide accuracy information of sensors measurements.
\item Provide an explicit description of the process by which an observation was obtained.
\item Provide an executable process chain for deriving new data on demand.
\end{itemize}

In SensorML sensors are modeled as processes that convert real phenomena to data. Processes take inputs, process them and result in one or several outputs. For example, a process can take a measurement generated by a thermistor which might be a resistance value and transform it to a celcius value. Processes can also be connected together in process chains to allow the sensor data to be processed by several processes sequentially. These processes and process chains are modeled in SensorML as shown in the  below SensorML Conceptual Model \ref{fig:SMLConceptualModel}.

\begin{figure}[h]
\centering
\framebox{\scalebox{0.65}{\includegraphics{SMLConceptualModel.png}}}
\caption{SensorML Conceptual Model}\label{fig:SMLConceptualModel}
\end{figure}

Processes in SensorML derive from AbstractProcess with it's inputs, output, parameters and metaDataGroup properties. The inputs and outputs represent data connections to other processes. Parameters represent inputs other than sensor measurements or outputs of other processes. They are used to configure the process or provide other inputs that are not sensor measurements. For example a parameter might be the latency time of a measurement needed for a calculation. Finally each process provides sensor metadata as modeled by the MetaDataGroup.

SensorML divides processes into two types, physical and non-physical. The non-physical process is one where  location, position etc. is not relevant to the process. Physical processes provide physical information, such as location, which can be used in processing if required. The ProcessModel is a non-physical process while the Component is a physical process and contains extra properties for defining the needed spatial or temporal information. Both process types contain the processMethod property which points to the methodology for transforming inputs to outputs.

There is a similar division of the process chains into the non-physical ProcessChain and physical System. Both process chains can contain a mixture of ProcessModel and Component process if desired. Process chains have inputs and outputs which define the beginning and end of the processing chain. They are points of connection to other processes and process chains if needed. The connection property contained in the ProcessChain and System is a list of Link type. The Link type contains source and destination properties that reference inputs and outputs of the processes in the process chain thus connecting the processes.

The SensorML specification currently contains the Sensor Web Enablement (SWE) Common specification. This specification defines basic types and data encodings used by the SWE specifications. It describes both primitive data types as well as aggregate types such as arrays and records. There are a large number of types defined in this specification, too many to show here. Below is a Conceptual Model of some of the simple data types.

\begin{figure}[h]
\centering
\framebox{\scalebox{0.65}{\includegraphics{SWESimpleConceptualModel.png}}}
\caption{SWE Common Simple Data Types.}\label{fig:SWESimpleConceptualModel}
\end{figure}

The data types in \ref{fig:SWESimpleConceptualModel} are scalar primitive values. They are the basis for the inputs, outputs and parameters in SensorML processes. Quantity models a floating point number. Count models a counting value or integer.  Boolean models a property that can be true or false. These types contain a number of properties that further define the type. The uom property provides a unit of measurement that indicate how the value should be interpreted.  A constraint property allows a value range or a enumerated list of values to be defined for the type. The quality property allows for a measure of its quality, for example the confidence level of a value being correct expressed as a percentage. Aggregation of these primitive data types is also provided by several types including DataRecord and DataArray.

SensorML uses XML Linking Language (XLink) \footnote{http://www.w3.org/TR/xlink} to support hypertext referencing in SensorML XML documents using URNs and URIs. This allows SensorML models to be distributed over several documents. XLink attributes are also used to reference on-line unit definitions used to define a sensor measurement.

The XML encoding of the SensorML model is not described here. The OGC SWE specification uses XML Schema to define the allowable SensorML XML structure.It is noted here that sml is the XML namespace for SensorML and swe is the XML namespace for SWE Common.

\section{Sensor Observation Service}\label{SOSSec}
SensorML is envisioned to be used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SOSref}. The SOS defines a web service interface for the discovery of sensors systems and the retrieval of observations from sensor systems. It organizes related sensor observations into a collection called an observation offering.

The SOS specification uses four main entities to describe the SOS's operation. The entities are Data Consumers, Data Producers, OGC Catalogs and the SOS itself. Data Consumers are clients of the SOS that request sensor observations and sensor metadata. Data Providers are entities have to access to the sensor and have enough processing power to generate an XML encoding of the sensor's observations. These observations are sent over a network to the SOS. The OGC Catalog is used to find SOS instances. SOS instances act as access points to a Sensor Web Network and provide the Web Service for retrieval of sensor observations. The figure \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operation context.
\begin{figure}[h]
\framebox{\scalebox{0.65}{\includegraphics{SOSoperationContext.png}}}
\caption{SOS Operational Context}\label{fig:SOSoperationContext}
\end{figure}

The diagram above shows the core and transactional operations provided by an SOS. The core operations are GetCapabilities, DescribeSensor and GetObservations. These are mandatory operations used by Data Consumers to request sensor information.  The core operations perform the following function:
 \begin{itemize}
\item GetCapabilities  -    retrieve SOS observation offering information.
\item DescribeSensor -    retrieve detailed information about the sensors.
\item GetObservation -   obtain sensor observations.
\end{itemize}

The transactional operations are RegisterSensor and InsertObservation used by the Data Producer to provided sensor information and observations. The transactional operations supported are described in the following:
 \begin{itemize}
\item RegisterSensor - operation to register the Data Provider with the SOS.
\item InsertObservation - operation to update SOS with a sensor observation.
\end{itemize}

These operations are encoded in XML and are intended to be carried http  operations. The sensor observations contained in  GetObservation and InsertObservation operations are encoded according to the OGC O\&M Observations \cite{OMref}. Sensor systems metadata contained in DescribeSensor and RegisterSensor is defined in SensorML or TML.

The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification. The SOS Data Consumer and SOS Data Provider entities will referred to a DataConsumer and DataProvider respectively in this thesis when describing the SenseTile Web Service.

The OGC Catalog Service is not explored in this thesis. The SOS Operation Context diagram shows a discovery process involving the  Data Consumer and a OGC Catalog Service (CS-W) catalog. The Data Consumer uses a CS-W catalog to find SOS instances. The GetCapabilites are then used toward each of the SOS instances returned to find sensor systems that are of interest to the Data Consumer. The CS-W catalog is covered in a separate specification \cite{OGCcatref}. This not covered in any real detail in the SOS specification.

There several optional SOS operations not covered in this thesis. The SOS refers to these as enhanced operations. They are GetResult,
GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest, DescribeObservationType, and DescribeResultModel.

\section{SenseTile}
A case study of the use of SensorML in sensor networks is based on the SenseTile System. SenseTile is a general purpose sensor system developed at UCD CASL \footnote{http://kind.ucd.ie/
products/opensource/SenseTile}. The SenseTile System is used to investigate issues arising with large scale sensor networking. SenseTile is made up of a sensor platform, a datastore, and a compute farm. Figure \ref{SenseTileDescription} shows a outline of this system. 
\begin{figure}[h]
\centering
\framebox{\scalebox{0.65}{\includegraphics{SenseTileDescription.png}}}
\caption{SenseTile System}\label{fig:SenseTileDescription}
\end{figure}
 The sensor platform is called the SenseTile Unit and is composed of one or more SenseTile Sensor Boards paired with a SenseTile Processor Unit. The Processor could be a PDA or small PC or any processing device that has a USB connection.  The Sensor Board has a USB connection used for communication with the Processor Unit. There are over a dozen sensors on the SenseTile Sensor Board including a thermistor and light sensor. Sensors are mounted on the Sensor Board and more sensors can be added using the USB connection.

This SenseTile unit is used here as case study on the use of SensorML, though SensorML can be used in other the parts of the SenseTile system. As the processor unit is a relatively powerful computer, providing a Web Service on the SenseTile Unit is feasible. The part SensorML plays in this Web Service is explored in the context of the SenseTile system.

Thermistor?

\section{UCD Sensor Board Driver}\label{SensorBoardDriverSec}

\section{VAST SensorML Engine}\label{VastSensorMLEngineSec}
The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source Java based SensorML Processing Engine. This software provides types that implement SensorML processes and process  chains. The SensorML Processing Engine parses SensorML models and instantiates the required objects to performed the intended processing. The code is found at \footnote{http://code.google.com/p/sensorml-data-processing} and is referred to as the VastSMLEngine in rest of the thesis. This code is dependent on an implementation of the SWE Common data types described in the section on SensorML \ref{SMLsection}. This code is found at \footnote{http://code.google.com/p/swe-common-data-framework}.

\section{BON} \label{BONsec}
Business Object Notation (BON) is a notation and a method for analysis and design of Object-Oriented (OO) systems. It supports a seamless transition from the design to the source code as it based on the same Object Oriented (OO) concepts supported by most OO languages. It is reversible as changes in the source code can be easily mapped back into the design and analysis models. It supports design by contract\ref{meyerref} where a function will guarantee a postcondition if a specified precondition is fulfilled by the caller of the function. BON support for design by contract allows for a formal description of a system to be developed. For a detailed description of BON see \cite{BONref}.
\chapter{Design}
\section{Overview}
This chapter describes the design of the SenseTile Web Service. Rather than define any new sensor models or XML formats for SenseTile it preferable to use the existing solutions if good enough for the task at hand. The OGC SWE framework as described in section\ref{SWESec} seems to cover all that is needed for modeling sensors and observations. Central to the design is a SensorML model of the SenseTile System. The section SensorML Model of SenseTile covers the details of this model. This model acts as both metadata about the SenseTile system and provides a description of how to process the sensor data for the SenseTile Web Service.
The Web Service design is based on the SOS specification and allows access to the SensorML model of SenseTile sensors as well as SenseTile sensor observations. Two SOS specification concepts as described in section \ref{SOSSec}, the Data Producer and the SOS itself, structure the design and are described in the SenseTile Web Service section.

\section{SensorML Model of SenseTile}\label{SenseTileModelSec}

A SensorML model of SenseTile is developed as part of this thesis. The SenseTile Sensor Board and its sensors is modeled in the most detailed. The model of the Sensor Board is placed in a context to examine SensorML process chaining where processing of Sensor Board data is performed in a separate SensorML process.  The SensorML Systems and Components that are used to model the system are shown in a  block diagram in figure \ref{fig:SensorML_SenseTile_System_comp}.

\begin{figure}[h]
\centering
\framebox{\scalebox{0.65}{\includegraphics{SensorML_SenseTile_System_comp.png}}}
\caption{SenseTile System SensorML Block Diagram}\label{fig:SensorML_SenseTile_System_comp}
\end{figure}

In the block diagram there are three SensorML Systems, the SenseTile System, the Sensor Board System and the Conversion System. The SenseTile System is used as container for the other two systems linking them in a process chain. The Sensor Board system contains metadata about the SenseTile Sensor Board and is a container for SensorML Components that model the Sensor Board's sensors.

The block diagram shows the physical phenomena measured by the sensors being feed as inputs to the Components in the Sensor Board System. The SensorBoard System outputs digital numbers based on the sensors measurement that have no physical meaning yet. The SenseTile System outputs the digital numbers directly and also connects this output to the input of the Conversion System. The digital numbers from the sensors are fed into the relevant conversion process that produce a meaningful value such as Celsius or Lumen. This allows the SenseTile Web Service provide the raw data from the sensors or the data converted to meaningful values to clients.

Only the Thermistor is modeled in this thesis. The  CelsiusConv Component converts the digital number form the Thermistor Component to a Celcius value. The same approach can applied to the other senors though for some components no conversion may be necessary.

A description of the SenseTile Systems and Components used in the SenseTile model is provided in the following sections.

\subsection{SenseTile System}
The SenseTile System is a SensorML System as described in section \ref{SMLsection} and is a container for the Sensor Board and Conversion Systems connecting their inputs and outputs as appropriate. The below XML fragment shows how SensorML encodes the containment of these Systems. 
\lstset{language=XML,basicstyle=\scriptsize,frame=single}
\begin{lstlisting}
<sml:components>
        <sml:ComponentList>
               <sml:component name="sensorBoardSystem" 
                                          xlink:href="SensorBoardSystem.xml">
               </sml:component>
               <sml:component name="convertorSystem"  
                                          xlink:href="ConverterSystem.xml">
               </sml:component>
        </sml:ComponentList>			   
 </sml:components>
\end{lstlisting}
The components and ComponentList elements contain a list of component tags. These component elements name the component which may be another SensorML System  Component. The component tag also has an XLink href used to reference the location of the SensorML definition of the component.  In this case the SensorML descriptions of the SensorBoard System and the Conversion System are in separate external files rather than in-line in the SenseTile System description. The use of XLink in SensorML is described in the SensorML section \ref{SMLsection}.

The SenseTile System connects up the inputs and outputs of the SensorBoard and Conversion Systems using SensorML connections. The below SensorML fragment shows how the temperature output from the SensorBoard System is connected to the input of the Conversion System.

\begin{lstlisting}
<sml:connections>
   <sml:ConnectionList>
       <sml:connection name="ambientTemperature">
          <sml:Link>
              <sml:source ref="this/inputs/ambientTemperature"/>
              <sml:destination ref="sensorBoardSystem/inputs/ambientTemperatureInput"/>
          </sml:Link>
   </sml:connection>
   <sml:connection name="temperatureDN">
          <sml:Link>
              <sml:source ref="sensorBoardSystem/outputs/temperatureDNOutput"/>
              <sml:destination ref="this/outputs/temperatureDNOutput"/>
          </sml:Link>
   </sml:connection>
   <sml:connection name="convertToTemperature">
          <sml:Link>
              <sml:source ref="sensorBoardSystem/outputs/temperatureDNOutput"/>
              <sml:destination ref="convertorSystem/inputs/celciusConvInput"/>
          </sml:Link>
   </sml:connection>
   <sml:connection name="outputTemperature">
          <sml:Link>
              <sml:source ref="convertorSystem/outputs/celciusConvOutput"/>
              <sml:destination ref="this/outputs/temperatureOutput"/>
          </sml:Link>
   </sml:connection>
</sml:ConnectionList>
\end{lstlisting}

The connection name attribute is used here to give a meaningful name to the connection. The Link type, as described in \ref{SMLsection}, is encoded with the Link tag. The contained source and destination elements contain a ref that indicates the components output and input to use. Only one connection is shown here but it is contained in a elements that allow a sequence of connection elements.

\subsection{SensorBoard System}

This System models the Sensor Board and contains meta data about the Sensor Board. It connects up the inputs and outputs the sensors.

\subsection{Thermistor Component}
The Thermistor on the SenseTile Sensor board is the Texas Instruments TMP175 Digital Temperature Sensor and is modeled with a SensorML Component. This is an indivisible process and can have a different location information than the Sensor Board if required. The SensorML Component allows the capabilities of the Thermistor to be described. For example the TMP175 is  specified for operation over a temperature range of −40°C to +125°C. This is captured in the following SensorML:

\begin{lstlisting}
<swe:field name="TemperatureRange" 
           xlink:arcrole="urn:ogc:def:property:dynamicRange">
          <swe:QuantityRange definition="urn:ogc:def:property:temperature">
                   <swe:uom code="cel" /> 
                          <swe:value>-40 125/swe:value> 
           </swe:QuantityRange>
</swe:field>
\end{lstlisting}

The SensorML Thermistor temperature input is modeled using a Quantity value without any units as is a measured physical phenomena. Its output is a digital number from the sensor. The range is constrained to the allowed value range from the Thermistor. The following fragment shows how the input and outputs for the Thermistor look in SensorML:

\begin{lstlisting}
<sml:inputs>
    <sml:InputList>
         <sml:input name="thermistorInput">
             <swe:Quantity definition="urn:ogc:def:phenomenon:temperature">
         </sml:input>
    </sml:InputList>
</sml:inputs>

<sml:outputs>
    <sml:OutputList>
          <sml:output name="thermistorOutput">
               <swe:Count>
	        <swe:constraint>
                         <swe:AllowedValue id="outputRange">
                              <swe:interval>-880 2032</swe:interval>
                         </swe:AllowedValue>
                     </swe:constraint>
                </swe:Count>
           </sml:output>
     </sml:OutputList>
 </sml:outputs>
\end{lstlisting}

\subsection{Converter System}
The Converter System is a container for the processes that convert the sensor data to a meaningful value. It this case it connects up the inputs and outputs the CelsiusConv Component. 
\subsection{Celsius Converter Component}

The SensorML Celsius Converter SensorML Component  models a process that performs a conversion of the Thermistors digital number output to a real quantity celsius value. It similar to the Thermistor SensorML description but its inputs and outputs are different. Its process method performs a simple calculation to perform the Celsius conversion.

A Sensor process method is also developed. The process method describes the conversion algorithm and points to an implementation class. Process Methods are not implemented by the Vast Lib so it not used in the running system. Instead the method tag provides a URN that is used to lookup the implementation class.

\section{SenseTile Web Service}
\subsection{Overview}

The SenseTile Web Service is based on the Sensor Observation Service (SOS) as described previously. The two main entities in the SOS Specification, the Sensor Data Provider and the SOS,  are used to structure the SenseTile Web Service. These entities are implemented as seperate software components that together provide the service. The Sensor Data Provider component will be referred to as the DataProvider in this design.

The two components are run as seperate processes on the SenseTile processor unit to allow a more flexible network architecture.  This network architecture is described in more detail in section \ref{subsec:SenseTile WebService Network Architecture}. The basic system structure is shown in figure \ref {fig:Deployment_sensetile} .

\begin{figure}[h]
\centering
\scalebox{0.65}{\includegraphics{Deployment_sensetile.png}}
\caption{SenseTile Web Service}\label{fig:Deployment_sensetile}
\end{figure}

The DataProvider parses a SensorML description of the SenseTile System, reads sensor data form the Sensor Board and provides the sensor data as O\&M Observations to the SOS. The SenseTile SOS will implement the SOS core operations to allow a client to get these observations. The following sections describe the design of the DataProvider and SOS components as well as the scenarios they support.

\newpage
\subsection{Supported Scenarios}
This section provides UML sequence diagrams to show the how the DataProducer and SOS interact to provide the sensor to the DataConsumer client.

\subsubsection{Generate Observations from Sensor Data}
The DataProducer accesses data from the sensor board 
\begin{figure}[h]
\centering
\framebox{\scalebox{0.65}{\includegraphics{InsertObs.png}}}
\caption{Generate Observation}\label{fig:InsertObs}
\end{figure}
 \begin{enumerate}
\item -
\item -
\item -
\end{enumerate}

\subsubsection{Get Sensor Description from SOS}
\begin{figure}[h]
\centering
\framebox{\scalebox{0.65}{\includegraphics{GetMeta.png}}}
\caption{Find SOS sensor}\label{fig:GetMeta}
\end{figure}
 \begin{enumerate}
\item -
\item -
\end{enumerate}
\subsubsection{Get  Sensor Observation from SOS}
\begin{figure}[h]
\centering
\framebox{\scalebox{0.65}{\includegraphics{GetObs.png}}}
\caption{Get Observation from Sensor}\label{fig:GetObs}
\end{figure}
 \begin{enumerate}
\item -
\end{enumerate}

\subsection{SenseTile WebService Network Architecture}
\label{subsec:SenseTile WebService Network Architecture}
In the SenseTile Web Service the SOS acts as a gateway to the SenseTile Sensor Network. This prototype is designed with the network architecture as shown in fig \ref{fig:Deployment_network}. A client accesses multiple SOS instances to get observations from the sensor network. An SOS can support several DataProducers and hence would not to be run on every SenseTile Unit.

\begin{figure}[h]
\scalebox{0.65}{\includegraphics{Deployment_sensetile_network.png}}
\caption{SenseTile Sensor Network Architecture}\label{fig:Deployment_network}
\end{figure}
The SOS Specification envisioned that the SOS is run on a large external server as shown in fig \ref{fig:Deployment_ext_sos}. The DataProducers update this external server with observations and SOS Data Consumer access this large SOS.
\begin{figure}[h]
\centering
\scalebox{0.65}{\includegraphics{Deployment_sensetile_ext_sos.png}}
\caption{External SOS}\label{fig:Deployment_ext_sos}
\end{figure}

The SOS Specification describes the use of a Proxy SOS eeicpr fix. This architecture could be used with the SenseTile Sensor Network if accessing many SOSs is not feasible for the clients.

\newpage
\subsection{DataProducer}\label{DataProducerHigh}
The DataProducer's role, as described previously, is to read sensor data from the Sensor Board and generate O\&M observations. These observations are then sent to the SOS instance. The processing of the sensor data to generate the observations is described by the SensorML models. The DataProducer uses the OGC Vast Library to instantiate objects that implement the required data processing chains based on the SensorML models. The BON static diagram \ref{fig:bon_static_diagam_producer.png}
 shows the main clusters and classes of the DataProvider.

The SMLEngine cluster contains classes that provide access to
the Vast Lib that is providing the SensorML Engine.  The Vast Lib implementation
of the SensorML Engine is hidden from the rest of the code thus.

The Sensors and Converters clusters contain classes that provide
implementation of process methods for SensorML components in the SensorBoard and
Converter SensorML Models respectively. These are referred to as sensor objects and
converter objects henceforth.

The DataProducer operates by repeating the below steps continuously.
 \begin{enumerate}
\item Read Sensor Board Data
\item Process Sensor Data
\item Create O\&M Observations
\end{enumerate}
The steps above are detailed the in following sections with reference
to the BON static diagram \ref{fig:bon_static_diagam_producer.png}.

\subsubsection {1. Read Sensor Board Data}

The UCD SensorBoard Library is used by DataProducer to access
the data generated by the Sensor Board.
The library provides high level interface to the
the data from the sensor board. It reads data from the Sensor board and generates packets
containing the sensor data. These packets provide functionality
to access a particular sensors data such as the temperature
measurement generated by the thermistor. 
This library is shown as the SensorBoard cluster in \ref{fig:bon_static_diagam_producer.png}
An instance of the class PacketInputStream to is used access the packets from the Sensor Board.

The DataProducer uses the Observer pattern to provide
packets to the SensorMLEngine for processing.
At start up the SenseTileSystem reads the list of sensor objects from the SMLEngine that
implement the Sensors interface. These objects provide an entry point
to the SensorML processing chain and can read data from the UCD LIB packets.
These sensor objects implement reading strategies for the type of data they read. 
For example the ThermistorSensor reads out the temperature value
but will average it before it sent to the SensorML Engine for processing.

\subsubsection {2. Process Sensor Data}
The PacketInputStream will request the SenseTileSystem to exexcute
the SensorML engine to process the sensor data. The SensorML Enigine
will execute the process chains described in the SenseTile SensorML model.
The sensor objects output will be passed to the converter objects. The
outputs of both object types are then available to generate observations.

\subsubsection {3. Create O\&M Observations}

When the SensorML Engine is finished
and there are obsevations generated these
need to be sent to the SOS. An O\&M Onservation
XML string containing the data is sent to
the SOS instance. This is timestamped at
this stage.
 containing the output from
the SensorML processes and send the observations
to the SOS.
-register offerings with the SOS. In this design
each system is an offering.
Need to generate an observation model?

The DataProducer was developed as a SensorML based software. It parses the SenseTile SensorML model to instantiate the code to execute the processes described. This required two aspects to be developed, a framework to run the processes and classes that implemented the components.

\begin{figure}[h]
\centering
\scalebox{0.65}{\includegraphics{bon_static_diagam_producer.png}}
\caption{DataProducer Static Diagram}\label{fig:bon_static_diagam_producer.png}
\end{figure}

\newpage
\subsection {SOS}

The SOS component of the SenseTile Web Service provides a simplified OGC SOS Web interface to access the observations from the DataProducer. The below is a static diagram showing the important clusters and classes in the SOS design.

 \begin{figure}[h!]
\centering
\scalebox{0.65}{\includegraphics{bon_static_diagam_provider.png}}
\caption{SOS Static Diagram}\label{fig:bon_static_diagam_provider.png}
\end{figure}
The SOS is the main cluster and the Service cluster contains the classes that provide the Web Service interface.

The DataProvider maintains a list of observation offerings provided by the SOS using the ObservationOffering class. When a new sensor type is registered a new ObservationOffering is created. If a DataProvider registers and provides an offering that already exists then just a new Sensor instance is added to the existing ObservationOffering.

The Sensor class contains a list of Observations that hold the observations from the DataProvider. The last 50 observations are maintained.

The ObservationService class provides the interface for the DataProducer to update with observations.

The SenseTile Service provides Web Service interface to be used by clients to access the sensor observations.

The SOS specification provides a very rich filtering feature that is too complex for this prototype. The operations are developed in a way to allow the sensor metadata and observations to be retrieved for a specific sensor board. The latest observation or all the stored observations can be retrieved.

\chapter{Implementation}
\section{Overview}
The SenseTile Web Service is written in Java. As described in the high level design there are two components, the DataProducer and the SOS, that are developed to provide this service. The two components are run in separate processes. The communication between the DataProducer and the SOS uses Java RMI. RMI is used due to the SOS and DataProducer
processes are running in the SenseTile Network and communicate over a LAN. This communication protocol has lower overhead than a Web Service so efficient communication between the DataProducer and SOS is achieved. Though it would not be hard to change to use a Web Service as a Java Interface is used to hide the communication mechanism from the main application code. Clients, SOS DataConsumers, access the SOS using a SOAP based Web Service interface.

The SOS specification describes an implementation using HTTP carry complete operation information encoded in XML including the operations name and parameters. The SenseTile WebService takes a different approach. The operation is defined in the RMI interface and the Web Service Interface while the parameters are encoded in XML thus need parsing.

The following sections describe the DataProducer and SOS implementation.
\section{DataProducer}
The DataProducer, described in section \ref{DataProducerHigh}, accesses the sensor board data and inserts observations into the SOS. Two libraries are used to achieve this: the SensorBoardDriver and the VastSensorMLEngine.

The SensorBoardDriver is described previously in section \ref{SensorBoardDriverSec}. There is also a SensorBoard Simulator library developed by UCS CASL researchers. It produces sensor board packets that can be feed into the DataProvider. This was used during the development of the DataProvider to facilitate testing.

The VastSensorMLEngine, described in section \ref{VastSensorMLEngineSec}, is used in conjunction with the SenseTile SensorML Model (section \ref{SenseTileModelSec}). The SensorML is parsed by the VastSensorMLEngine into a DOM\footnote{http://www.w3.org/DOM/} tree. To perform the processing it propagates an execute method to the instantiated processes. A ProcessMap.xml file maps SensorML ProcessModel and Component method  URNs  (see section \ref{SMLsection}) to Java classes that provide the implementation of the processing to be performed.

There is no documentation that I can find for the VastSensorMLEngine. There are some reference implementations of SWE services pointed to from the OGC SWE website \footnote{http://www.ogcnetwork.net/SWE} that provide some information how to use it. The SenseTile SensorML model was not parsed initially by the VastSensorMLEngine. As it opensource the code is available and after a few small changes the model parsed. The VastSensorMLEngine processes the model fine and was very stable during testing.

\section{SOS}
The SOS is the provider of the Web Service interface and is built on Apache Axis2 \footnote{http://ws.apache.org/axis2}. 
Axis2 is described as a "Web Services / SOAP / WSDL engine"  with the Apache Axis2/Java implementation used in this thesis. The Spring framework\footnote{http://www.springsource.org/} was also used. Using these libraries a light Web Service is built to run on the reasonably powerful SenseTile Processor Unit. Axis2 is run standalone here though it can be on with any Servlet engine if a more powerful Web Service is needed.

Developing a Web Service is straight forward with Axis2 and it supports several approaches. The supported POJOs (Plain Old Java Objects) approach was used for the SenseTile Web Service allowing the code to be reused on a different framework if AXIS2 did not work well. Axis2 supports the The Spring framework which is used in the SenseTile WebService for it's dependency injection functionality. A SprinitInit class is needed for Axis2 to load up the Spring framework.

Axis2 provides a tool that automatically generates WSDL from the running Web Service. Another tool allows client side code to be generated from the WSDL easing the development of the Web Service clients.

JAXB \footnote{http://java.sun.com/developer/technicalArticles/WebServices/jaxb} was used for SensorML XML to Java binding in the Web Service. JAXB is able to generate a default binding from the complex SWE schema. Some name collisions were corrected with a JAXB customization file found on the SensorML newsgroups. However the default binding was not fully successful for the System type. It input, output and parameter properties were missing and the inheritance hierarchy did not provide them. A small change was needed to add these properties and then everything worked fine with marshaling and unmarshaling SensorML Systems.

\chapter{BON to SensorML mapping}
\section{Overview}
In this chapter a BON(see section\ref{BONsec}) to SensorML( see section \ref{SMLsection}) mapping is described. One motivation for this mapping is take advantage of formal software development processes described in \cite{Kiniryref}. This is a verification based process that refines BON to JML and then to Java. Robust software can thus be developed. Another motivation is that it XML is verbose and a more compact description of a SensorML model as facilitated by BON is naturally preferred. Lastly SensorML is complex and it is difficult to develop correct SenorML models based only on using the SensorML XML Schema definition. How a BON specification of SensorML can help here is described later in the chapter.

\section{SensorML BON Specification}
The BON specification of SensorML described is very close conceptually to the SensorML model shown in section \ref{SMLsection}.  The BON SensorML model is shown in figure \ref{fig:bonSensoMLModel}. 
\begin{figure}[h]
\centering
\scalebox{0.65}{\includegraphics{bonSensoMLModel.png}}
\caption{BON SensorML Model}\label{fig:bonSensoMLModel}
\end{figure}
The inheritance heirarhcy is changed slightly with System inheriting from ProcessChain and Component inheriting from ProcessModel. A new abstraction called Port is added. Port contains all the properties of SensorML input, output and parameter as they all have the same properties. A PortType class that ensures legal port types of Input, Output and Parameter is also added to the SenorML BON model.

A BON specification SWE Common data types Count and Quantity are also described along with needed dependent classes. The abstract class DataComponent is the base type for these. There are a many other types but only Count and Quantity are developed as they were the ones used in the SenseTile SensorML model.

The BON specification developed in this thesis does not address the issue of the SensorML metadata and the MetaDataGroup is not specified. This information is for descriptive purposes only and is not used for processing sensor data. BON adds no value here. A different way to add this information to SensorML sensor descriptions is suggested to be used.
AssociationAttributeGroup
SensorML
Member
The complete BON SensorML specification listing can be found in appendix \ref{appenBSMLBONSpec}.

\section{SensorML BON to XML Mapping}
SensorML is essential XML defined by XML Schema with its main purpose being interopeability. Therefore a mapping SensorML BON specification to the SensorML XML is needed. The BON types map closely to SensorML complex types and so a relatively simple scheme is needed.
Effective BON classes map to XML elements. Features in the class map either to attributes of the XML element or a contained XML element. Whether a feature is an attribute is indicated in the class as a comment using @SensorML to make this comment as a mapping instruction. Otherwise an XML element is assumed. The type names of the BON classes and features are the XML element or attribute names. Values. Case of names.

BON SEQUENCE types need a special rule for mapping to SensorML.

Connections rule.

\section{ProcessChain connections problem}
One of the major features of SensorML is the ablity to describe processing chains. The SensorML ProcessChain and System types have a connections property that allows process chaining as decribed in section \ref{SMLref}. Strings are used to describe the connection. An example of this is shown in listing \ref{connectionListing} shows one of the connections from the SenseTile System as described in section \ref{SenseTileModelSec}
\begin{lstlisting}[label=connectionListing,caption=SenseTile System connectionconvertToTemperature]
   <sml:connection name="convertToTemperature">
          <sml:Link>
              <sml:source ref="sensorBoardSystem/outputs/temperatureDNOutput"/>
              <sml:destination ref="convertorSystem/inputs/celciusConvInput"/>
          </sml:Link>
   </sml:connection>

\end{lstlisting}

The source and destination reference strings need to be parsed to find out the inputs and outputs to connect. This is very error prone and the ability to check that connections are correct to a least some degree is desirable. A BON specification can enforce some level of correctness in setting up the connections using preconditions (ensure). The source reference must be either an input of the process chain itself or an output of one of the contained components. In the BON ProcessChain class a feature AddLink checks that Links are legal before adding the Link to the connections feature. This is shown in the BON listing below.

\begin{lstlisting}[basicstyle=\scriptsize,showspaces=false,showstringspaces=false, caption=BON ProcessChaing]  
effective class ProcessChain
     indexing
     about:        "Process formed by chaining sub-processes.";
     title:        "ProcessChain.";
     author:       "Ciaran Palmer.";
     copyright:    "none.";
     organisation: "School of Computer Science and Informatics, UCD.";
     date:         "2010/04/01.";
     version:      "Revision: 1.00.";

     inherit AbstractProcess
     feature

     make
       ->componentsIn:AbstractProcess
       require
         componentsIn/=Void;         
       ensure
         delta{components};
         components = componentsIn;
       end

     -- Collection of subprocesses that can be chained using connections      
     components:SEQUENCE[AbstractProcess]

     -- Links between processes
     connections:SEQUENCE[Link]

     -- Add a link with check that Link is legal
     AddLink
        ->linkIn:Link
      require
          linkIn /= Void;

           -- check Link source is legal
          (exists i:INTEGER such_that linkIn.source = Current.inputs[i]) or\
          \(exists i:INTEGER and j:INTEGER such_that linkIn.source =\
          \components[j].outputs[i]);

          -- check Link destination is legal
          (exists i:INTEGER such_that linkIn.destination = Current.outputs[i]) or\
          \(exists i:INTEGER and j:INTEGER such_that linkIn.destination =\
          \components[j].inputs[i]);

        ensure
           delta{connections};
          (exists i:INTEGER such_that connections[i] linkIn);

      invariant
        components /= Void;

  end --ProcessChain
 \end{lstlisting}

\section{SenseTile Thermistor BON SensorML Specification }
Thermistor can be generated by the following BON class.

\chapter{Results}

A  SensorML Model description of a SenseTile Sensor Board and Thermistor was developed. It was
run on a SensorML engine. Appendix shows the model developed.

Developed lightweight Web Service to explore the use of SensorML in Sensor Networks.
running prototype tested with UCD SensorBoard Simulator code.

Tested on a low power Atom PC that is used as SenseTile Processing Unit.

BON mapping developed for main SensorML types. A BON description of SenseTile SensorML Model
was developed based on the mapping.

\chapter{ Conclusions and Future Work}
This thesis analyses the role of SensorML and SOS specifications in Sensor Systems. The UCD CASL SenseTile System is used as a case study for the use of these specifications. A SensorML model of SenseTile and a SenseTile Web Service prototype.

SensorML address the issues of interoperabilty of Sensor Systems
rich description of a sensor systems. Hard to think of any system that could no be described by SensorML. Weather systems to Geloloaction services. 
Accurate elements. Processing data. Metadata.
The supporting SWE Common data types are many data types.
SensorML Frameworks.
Precisely define its data.
It also is bases for developement for Sensor Ontology work.\footnote{http://www.calebgoodwin.com/SEMANTIC.html}
SensorML is relatively intuitive but it is difficult to know if one has developed a "good" SensorML model. There are examples and tutorials to help.

The SenseTile model uses a hierarachy to model the SenseTile. It seems repetitive to have keep defining the same inputs for each contained component. Maybe this is a flaw in the SenseTile model developed.

SensorML is defined by a heavy XML Schema. This XML Schema along with the other XML Schema for the SOS and dependent specifications are heavy indeed. Achieving a Java binding was difficult. JIBX would not generate a default XML binding. JAXB could with some help. There are many layers of abstraction in these XML Schema. This reflects the that the specifications cover a very wide range of needs of a senor system.

The idea of automatically obtaining a SensorML model and executing it depends on the SensorML Engine being used. Only one Java implementation can be pointed. In the OGC examples the Java implementations of the

Vast Lib use as a references only.
Had to update library code to get it to work with the SenseTile schema even though the schema
passed the validation test tool from the site. It uses DOM but XML binding is probably a better approach.

The SWE SOS specification describes the operations for accessing sensor observations and it's concepts were used to develop the SenseTile Web Service. It uses SensorML provide sensor metadata to clients and the O\&M specification to encode observations.
Implement all functionality would be difficult. Maybe too heavy for SenseTile System. Simple querires. Or better a push system to send the data to the client as it is generated could be a better approach here.

SWE not fully web services as W3C might define. Not SOAP. HTTP envelope
flexible. There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
what has been achieve

SOS on the Processor Board. Is this right? RMI only useful if SenseTile Network is a lan.

BON

Working in XML is error prone. SensorML connections type was a good example of where a graphical tool or another language, like BON,  that can be parsed should be used to check the connections. Running the model through the VastSensorMLEngine without a crash was only way other than by eye to see that the connections were set up correctly.

further work.
---------------
SensorML sensor accuracy.
Complete BON mapping and parser.
Streaming the data. See SWE Architecture dpcument for a description of one approach.
Other technology approaches such as . Other ways such a XQUERY? XML techs?

\newpage
\begin{thebibliography}{99}
\bibitem{OSWAref} Chu, Kobialka,  Durnota, and  Buyya, Open Sensor Web Architecture: Core Services
\bibitem{SMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007
\bibitem{SOSref}Open Geospatial Consortium Inc.,  Sensor Observation Service, 2007
\bibitem{Kiniryref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
\bibitem{OMref}Open Geospatial Consortium Inc., Observations and Measurements – Part 1 - Observation schema, 2007
\bibitem{SWEArchref}Open Geospatial Consortium Inc., OGC Sensor Web Enablement Architecture, 2008
\bibitem{SMLTutorialref}Open Geospatial Consortium Inc., Using SensorML to describe a
Complete Weather Station , 2006
\bibitem{ProcessTutorialref}Open Geospatial Consortium Inc.,
\bibitem{OGCcatref}Open Geospatial Consortium Inc., OpenGIS® Catalogue Services Specification, 2007
\bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
\bibitem{Meyerref}Bertrand Meyer., Object-Oriented Software Construction, 1997

\end{thebibliography}

\appendix
\chapter{Appendix: SenseTile SensorML Model}
\section{SenseTile SensorML System}
\section{SenseTile SensorML SensorBoardSystem}
\section{SenseTile SensorML ConvetorSystem}
\section{SenseTile SensorML Thermistor}
\lstinputlisting{Thermistor.xml}
\section{SenseTile SensorML CelciusConvertor}

\chapter{Appendix: BON to SensorML Mapping}\label{appenB}
\section {SensorML BON Specifcation}\label{appenBSMLBONSpec}

\lstset{basicstyle=\scriptsize,showspaces=false,showstringspaces=false}
\lstinputlisting{SensorML_formal.bon}
\section {BON to SensorML mapping tables} \label{appenBSMLBONTables}
\lstset{frame=none}
\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class AbstractProcess  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class System  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class C  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class Link  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class ProcessChain  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class ProcessModel  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class Port  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class Quantity  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\begin{table}[!th]
\centering
\begin{tabular}{|l|l|}
\hline
BON features & SensorML XML tags\\
\hline
   deferred class Count  & not mapped\\
\hline     
     inputs:SEQUENCE[Input] & \begin{lstlisting}
<sml:inputs><sml:inputlist>
SEQUENCE[Input]
<\sml:InputList><\sml:inputs>\end{lstlisting}\\

\hline 
     outputs:SEQUENCE[Output] & \begin{lstlisting}
<sml:outputs><sml:outputlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\
\hline
     parameters:SEQUENCE[Parameter] & \begin{lstlisting}
<sml:parameters><sml:parametertlist>
SEQUENCE[Input]
<\sml:outputList><\sml:outputs>\end{lstlisting}\\

\hline                 
     metaDataGroup:MetaDataGroup &  \begin{lstlisting}
<\sml:metaDataGroup><\sml:metaDataGroup>\end{lstlisting}\\
 \hline

\end{tabular}
\caption{BON to SensorML AbstractProcess Mapping table}\label{table:bon_sml_example}
\label{ex:table}
\end{table}

\label{endpage}
\end{document}

\end{article}