### 2010-08-24 papers/AgileFormality/paper.tex
Revised sections 1, 5, and 6.

\documentclass[english]{lni}
\usepackage{latin1}

%\usepackage{times}
\usepackage{ifpdf}
%\usepackage{a4wide}
%\usepackage{pdfsync}

\ifpdf 
  \usepackage[pdftex]{graphicx} 
\else 
  \usepackage{graphicx} 
\fi

\usepackage{xspace} 
%\usepackage{tabularx} 
%\usepackage{epsfig}
%\usepackage{amsmath} 
%\usepackage{amsfonts} 
%\usepackage{amssymb}
%\usepackage{eucal} 
%\usepackage{stmaryrd} 
\usepackage{float}
\usepackage{listings}

\input{jml-listings}
\lstset{language=[JML]Java,xleftmargin=20pt,xrightmargin=20pt}

\ifpdf 
  \usepackage{epstopdf}
  \usepackage[pdftex,bookmarks=false,a4paper=true,plainpages=false,naturalnames=true,
  colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue,
  pdfauthor="Vieri del Bianco and Joseph R. Kiniry and Dragan
  Stosic"]{hyperref} 
\else
  \usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

\newcommand{\tablesize}{\footnotesize} 
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace} 
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
\newcommand{\myhreffootnote}[3]{\myhref{#1}{#2}\footnote{#3\myhref{#1}{#1}}}
\newcommand{\lil}[1]{\texttt{\lstinline|#1|}}

\newcommand{\notev}[1]{\xspace\marginpar{\scriptsize\textsf{Vieri:}#1}}
\newcommand{\notej}[1]{\xspace$\frac{\varocircle}{\textsf{jk}}$\marginpar{\scriptsize\textsf{Joe:}#1}}
\newcommand{\noted}[1]{\xspace$\textcolor{red}{\dagger^\textsf{dragan}}$\marginpar{\scriptsize\textsf{Dragan:}#1}}

\newcommand{\todo}[1]{\texttt{\textbf{TODO:} #1}}

\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\SB}{\emph{Sensor Board}\xspace} 
\newcommand{\STSB}{\ST \SB\xspace} 
\newcommand{\STSBs}{\ST \emph{Sensor Boards}\xspace}
\newcommand{\STPU}{\ST \emph{Processor Unit}\xspace}
\newcommand{\STPUs}{\ST \emph{Processor Units}\xspace}
\newcommand{\STU}{\ST \emph{Unit}\xspace} 
\newcommand{\STUs}{\ST \emph{Units}\xspace}
\newcommand{\mole}{``\emph{Mole}''\xspace}

\newcommand{\simulator}{\STSB \emph{Simulator}\xspace}
\newcommand{\datastore}{\ST Scientific Datastore\xspace}
\newcommand{\computefarm}{The \ST Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \ST Software and Data Compute Server Farm\xspace} 
\newcommand{\sensorfarm}{The \STSBs\xspace}

\begin{document}

\title{Agile Formality: A \mole of\\Software Engineering Practices}

\author{Vieri del Bianco and Dragan Stosic\\
  University College Dublin, Ireland\\
  \texttt{vieri.delbianco@ucd.ie} and \texttt{dragan.stosic@gmail.com}\\
  \\
  Joseph R.~Kiniry \\
  IT University of Copenhagen, Denmark\\
  \texttt{kiniry@itu.dk}
}

\maketitle

\begin{abstract}

  Members of the agile programming and formal methods communities do
  not always see eye-to-eye.  These two communities often do not talk
  to or learn from each other.  Only recently, as highlighted by the
  September 2009 issue of IEEE Software, the IFIP workshop on
  balancing agility and formalism in software engineering, and the
  first edition of the international workshop for formal methods and
  agile methods, ideas from the two communities begun synthesize.
  While the problem-solving approaches and psychological attitudes of
  members of the two communities differ widely, we exploit this clash
  of viewpoints, creating a new development processes that actually
  blends, rather than mashes together, best practices from the two
  worlds.  This paper summarizes our process and a supporting complex
  case study, showing that it is not only possible, but tasty, to
  combine the ``chili pepper'' of formal methods and the ``chocolate''
  of agile programming, thus producing a tasty \mole (as in the
  highly-spiced Mexican sauce) of software engineering practices.

\end{abstract}

\section{Introduction}
\label{sec:introduction}

% agile and formal difficult to blend: mind attitudes
Agile and formal development methodologies usually do not blend
together well.  This is because of several reasons, the most important
of which is often characterized as a radical difference in
psychological attitudes about software development.

% FM
Formal methodologists often favour an in-depth, think-first approach,
where the problem is understood, formalized, and solved; usually, but
not always, adopting a waterfall-style of development.  Once a
formalization is developed, development and verification proceed in an
automated and interactive fashion.  Consequently, projects that use
formal methodologies (FMs) typically focus on critical systems with
fixed requirements and somewhat flexible deadlines (``We'll ship it
when it is right.'').

% AM
Agile methodologists favour an highly incremental and iterative
approach, specifically tailored to cope with changing requirements and
precise deadlines.  In projects that use agile methodologies (AMs), it
is often the case that the problem is only partially understood, and
there is a focus only on the aspects that must be implemented in the
current development iteration.  The solution initially developed is
usually not optimal, but as it is refined through continuous code
refactorings, its quality and validity improve.  Test suites are used
for several purposes, the most important of which are system
verification, system documentation, and requirement specification, and
they are strictly handmade.  AMs are typically used in development
settings where changing requirements and rapid delivery of the product
are paramount.

% problems would benefit form agile and formal mix, characterize the problem
The two worlds seem irreconcilable.  Nevertheless, problems exist that
would greatly benefit from both of them~\cite{Black2009}.  The problem
class in which we have particular interest have \emph{unstable
  requirements}, are \emph{constrained by deadlines that cannot be
  postponed}, but also have \emph{subsystems that must be formally
  specified and verified}.

\subsection{Terminology}
\label{sec:terminology}

% IEEE terminology
In this paper the IEEE 610~\cite{Jay1990} standard terminology for
validation and verification is used.  Paraphrasing the standard,
\emph{verification} is the process of evaluating software to determine
whether the products of a given development phase satisfy the
conditions imposed, according to the specifications (``you build the
system in the right way''); \emph{validation} is the process of
evaluating software to determine whether it satisfies specified
requirements ensuring it meets the user's needs (``you build the right
system'').

% FM terminology
Within the FMs community, verification and validation have different
meanings than in the IEEE 610 standard.  Loosely \emph{verification}
(in the paper \emph{formal verification}) means the use of formal
methods to formally, statically verify that a system conforms to its
specification; \emph{validation} (in the paper \emph{formal
  validation}) means the use of traditional software engineering
practices to check, statically or dynamically, that a system conforms
to its specification (whether in the form of tests, requirements,
etc.) by hand or via execution under some execution scenarios.  In the
present paper both perspectives are respected and acknowledged.

% tests terminology
Tests are usually classified by dichotomies: unit vs. functional,
black-box vs. white-box, testing vs. design, tester vs. coder.  Yet
testing does not always fit into these dichotomies, especially in the
case of agile testing approaches~\cite{Beck2007}.  The black-box
vs. white-box dichotomy is particularly problematic, since box
boundaries change one's perspective.  In this paper, functional tests
are always black-box tests, coverage tests are always white-box tests,
generated tests are always black-box tests (the box is the module
specified); other strict rules do not hold.

% development terminology
Additionally, ``...development of a system,'' more precisely means
``...analysis, design, development, verification, formal validation of
a system.''

\subsection{Case Study}
\label{sec:case-study}

Proposing new practices with no supporting evidence is an
uninteresting proposition.  A complex case study is mandatory to
test-bench the proposed practices.  Our case study focuses on the
development of a device driver to control and communicate with an
embedded custom circuit board equipped with more than a dozen sensors.
The communication protocol with the board is asynchronous and
packet-based and the board is novel and custom-designed.  The protocol
is focused on controlling the board and reporting sensor measurements
and board status.

This software driver has been developed using our process while
satisfying seemingly contradictory requirements: (1)~the protocol
implementation must be thoroughly verified; (2)~parts of the protocol
must be formally specified; (3)~a board simulator must be designed,
implemented, and verified to match the specifications of, and actual
behaviour of, the physical circuit board; (4)~if board components
change during development, the associated formal specifications and
the simulator whose behavior reflects such changes must be updated
accordingly; (5)~there are strict software delivery deadlines.
Consequently, these process requirements constitute an environment
that benefits from both agile and formal approaches to software
development.

A novel development process has been drafted to cope with the
contrasting requirements of the case study, blending agile and formal
practices.  In the following, the development process and the practices
are summarized, and their successful application to the case study is
reported, yielding several promising results.

\subsection{Formal and Agile Integration}
\label{subsec:formal_and_agile_integration}

% what is needed, high-level
To enable the integration of the two worlds, the highly iterative and
incremental approach found in most of the AMs must be maintained, and
formal validation methodologies, traditionally used in a waterfall
development process (formally specify the system, implement the
system, validate the implementation against the specification) must
be adapted to a highly iterative one.

% what must be considered: TDD
The most common software verification practices in AMs must be
considered when trying to integrate FMs and AMs since they are a
fundamental part of the development process.  The most significant
verification related practice is Test Driven Development
(TDD)~\cite{Beck2003}.  TDD is a software development technique,
originally defined in the Extreme Programming (XP)~\cite{Beck2004}
methodology and is characterized by a process that focuses on writing
unit tests for a module before writing the code of said module.

% test driven kinds
The test driven approach relies on writing tests before
implementations.  It is applicable to a scall and large-scale cycles.
In the former case, very small scale TDD yields activity cycles as
short as a few minutes.  In the latter, requirements are immediately
translated into functional tests resulting in activity cycles as long
as a full delivery iteration, which is usually no longer than a couple
of months.  The test driven approach is a cornerstone of XP, but it is
so popular that has been adopted in many other AMs as well.  It is
also considered a good software development technique when used on its
own, as it pre-dates XP and AMs considerably, regardless of the
enclosing development process.

% previous related work
Some tentative attempts to reconcile FMs and AMs have been developed
by others.  The reoccuring theme of these attempts is that idea that
handmade test suites are no longer used, but instead are replaced by
different kinds of automated formal validation based on a system-level
formal specification.  Automated validation can be as simple as
enabling runtime checking of assertions or automatically generating a
test suite based on the system properties, or as complex as proving
the whole system's behaviour through formal verification.

% a quick analysis of the related works
Eleftherakis and Cowling in~\cite{Eleftherakis2003} propose XFun, an
iterative and incremental process to generate X-Machines models, an
extension of Finite State Machines.  The development process impose a
complete specification of the system, and its verification is done
exclusively using the tools provided by the X-Machines FM.  The
development process proposed by Herranz and
Moreno-Navarro~\cite{Herranz2003b} is quite similar.  They propose an
iterative process to model a system using SLAM-SL, an object-oriented
specification language supported by tools.  Some XP practices are
fully adopted or considered compatible, like pair-programming,
iterative and incremental development, and system refactoring; while
tests are completely replaced by the verification tools provided by
the SLAM suite.  A different FM, but similar approach, is found also
in the work of Suhaib~\cite{Suhaib2005} where the XFM methodology is
used.

% previous work: a common approach
As already mentioned, in all of these attempts, test suites are
automatically generated using the system formal specification, or are
completely replaced by formal verification, typically through the use
of theorem provers and/or static checkers.  The major precondition on
the use of these approaches is, at the minimum, a specification of all
the relevant aspects of the system under development.  Such a
specification enables simple verification techniques such as runtime
checking (via assertions and design by contract~\cite{Meyer1997} using
preconditions, postconditions and invariants) and the automatic
generation of test suites~\cite{Cheon2002,Cheon2004,Cheon2005}.  Via a
complete and sound system specification, supported by an appropriate
formal language (to specify the system model and its properties)
coupled with a (possibly restricted) programming language, one can
statically prove properties about the system using a variety of tools
and
techniques~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.

% previous work: why is not satisfying
The underlying idea of this approach is simple: what was once the
purpose of handwritten tests (verification, documentation, and
requirement specification) are now the responsibility of the formal
specifications, and a complete formal specification of the system
under development is needed to apply these methodologies.  The problem
is that this requirement is very difficult to fulfill since the effort
required to write a complete formal specification of the system in
most real-world complex cases is usually greater than writing a suite
of tests~\cite{Glass2002}.

If a complete formal specification is not available, all of the
previous methodologies share a common problem: the parts informally
specified are not verified at all.  These parts of the system are
consequently verified with traditional methods, but the connections
between development artifacts and related activities (code, tests,
design, refactoring) are neither detailed nor enforced.  This is
problematic, as the loops and relations between the different
development artifacts are the inner engine of many AMs, as they impose
an iterative and incremental pace to the development process.

% outsider
Liu takes a different approach wherein the SOFL
methodology~\cite{Liu2004} is merged into agile
processes~\cite{Liu2010}.  The SOFL methodology is based on a three
step specification approach: (1) informal specification, (2)
semi-formal specification, (3) formal specification.  The agile
adaptation consists mainly in introducing shorter loops between
activities and reducing the size of the set of formal specifications.
The formal specifications are used only to help understand ambiguous
statements in the semi-formal specification and are not maintained
while the system evolves.  Testing and inspections are based on the
semi-formal specification.  The formal specification are partial and
are only used for documentation purposes and then discarded---they are
not used at all in the verification of the system.

% what we want to achieve
In contrast, our objective is to formally specify only those parts of
the system that must be verified (in order to cope with constrained
resources and unstable requirements) and to develop and refine a
highly iterative and incremental development process that blends
formal and agile practices.

% how we want to achive it
The main open problem unaddressed in the literature is how to connect
development artifacts of the two worlds of FMs and AMs together.  To
answer this challenge we define activity cycles, similar to what is
found in AMs (especially in XP), to solve this problem in an highly
iterative development process.  We show how tests drive the formal
specification, how the formal specifications drive test and code
development, how handmade and automated tests coexist and support each
other, how the unspecified parts of the system are incrementally
specified.  That is, we show how to blend, and not just tack together,
formal validation and agile verification practices: a \mole of
verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly
  spiced Mexican sauce made chiefly from chili peppers (agile) and
  chocolate (formal), served with meat (working system).}.

\section{A Real and Complex Case Study: \\ Rapid Development in Small-scale Hardware Software Co-design}
\label{subsec:a_real_and_complex_case_study}

\mole practices are applied in the context of a real world case study that matches the problem requirements previously detailed: unstable requirements and development artifacts that need to be formally specified and validated.

\subsection{The UCD CASL SenseTile System}
\label{sec:ucd-casl-sensetile}

% SenseTile introduction
The \textbf{UCD CASL SenseTile System} is a large-scale, general-purpose sensor system developed at the University College Dublin in Dublin, Ireland.  
The facility provides a scalable and extensible infrastructure for performing in-depth investigation into both the specific and general issues in large-scale sensor networking. 
This system integrates a sensor platform, a datastore, and a compute farm.  
The sensor platform is a custom-designed but inexpensive sensor platform (called the \ST) paired with general-purpose small-scale compute nodes, including everything from low-power PDAs to powerful touchscreen-based portable computers.  
Besides containing over a dozen sensors packaged on the \ST itself, the board is expandable as well, as new sensors are added to it easily.

% case study: the sensor board development tasks
The case study is focused in building the sensor board and its software driver concurrently.  
Because of hard time constraints and the initial unavailability of the custom board, we must progress concurrently with all the development tasks, including: (1)~specification of the communication protocol; (2)~specification of the physical sensor board; (3)~development and fabrication of the physical board; (4)~development of the embedded software for the board; (5)~development of the communication protocol software driver of the board; and (6)~development of software simulators of the board.

% what we have to do, what we outsourced, and the dependencies between tasks
The development of the physical board, along with its embedded software (development tasks~2, 3, 4), is carried out by a third party under our guidance, thus it is not directly taken into account here. 
The specification of the communication protocol (development task~1) is a joint effort between ourselves and the third party, while the remaining tasks~(5, 6) are performed in isolation.  
Dependencies are straightforward: the specification of the communication protocol~(1) is the most stable element, but it still depends on the sensor board~(2, 3, 4): large changes in the latter affect the former.
Additionally, the driver and the simulator depends directly on the communication protocol.

% communication with external manufacturer
Communication and synchronization with the external manufacturer is frequent but not optimal.  
The main synchronization artifact is the protocol specification, which remains stable at high-level, but is changed frequently in the low-level details.  
A common domain language has been found and agreed upon, but the FMs and programming platforms differs widely.  
Consequently the informal specification of the communication protocol is the most reliable source of information.

\subsection{Related Methodologies}
\label{sec:relat-meth}

% how the problem is usually managed
There exist various approaches in literature addressing this kind of development constraints (rapid development in hardware software co-design), but all are focused on large-scale systems.

% hardware-software
The hardware-software formal co-design methodologies usually have several common development artifacts~\cite{Slomka2000, Hoffman2001} including a high-level specification of the system, a translation (refinement) of the high-level specification to low-level ones (hardware and software counterparts), the possibility to generate software code from the low-level specifications, a hardware simulator capable of simulating an hardware component based on hardware specification, the hardware component developed.

% device drivers and formal specification
When considering the development of the device and of its software driver as separate entities that possibly have to be developed concurrently, the existing approaches are similar to the ones seen in the case of hardware-software co-design~\cite{Valderrama1995,Siegmund2002,Ryzhyk2009}, all of which focus on a specification that aims to be as complete as possible.
A complete specification is neither feasible nor convenient in our case. 
The protocol specification must be enhanced incrementally and its complete specification cannot be provided.  
This makes the problem an ideal candidate to test our \mole of practices.

\subsection{The Chosen Formal Methodology: Formal Specifications with JML}

% formal methodology constraints
An appropriate FM must be chosen.  
It must be able to support a specification that is built incrementally, consequently formal methods that demand a complete specification must be avoided.
It must be able to automatically generate tests.
And finally the formal method must be complemented by tools that support the verification of system properties at runtime.

% the chosen formal methodology
The Java programming language and platform are chosen for the implementation, the Java Modeling Language (JML)~\cite{LeavensBakerRuby-Prelim} is used to write formal specifications, and JML2~\cite{JML2} tool suite is used as supporting tool~\cite{BurdyEtal05-STTT}.  
The JML2 tool suite includes runtime assertion checker~\cite{Cheon-Leavens02b} and a unit test generator~\cite{Cheon2002}.
A complete specification is not required to use JML2 tool suite effectively.

% JML introduction
JML is a rich behavioural interface specification language (BISL) and focuses on the modular specification of classes and objects~\cite{LeavensBakerRuby99}.  
JML includes standard specification constructs like assumptions, assertions, axioms, and pre- and postconditions, framing clauses, and invariants.
It also includes a rich model-based specification layer that includes mathematical models, data refinements, datagroups, and many other advanced object-oriented specification features~\cite{Chalin06}.  
Many tools, ranging from compilers to static checkers to full-functional verification systems support the JML language~\cite{BurdyEtal05-STTT}.

% using JML
JML is sometimes used in a Design-by-Contract style, where a specification is written from scratch, reusing existing APIs that have
specifications of their own, and then an implementation is written conforming to that specification~\cite{Meyer92b}.  
At other times an existing piece of implemented and tested code is annotated with specifications after-the-fact.

% JML tools
The two tools used most frequently to check the correctness of implementations are the JML Runtime Assertion Checker (RAC) and the
Extended Static Checker for Java, version 2 (ESC/Java2)~\cite{Cheon-Leavens02b,BurdyEtal05-STTT,KiniryCok04}.  
The former compiles executable assertions into runtime checks.  
The latter tool performs automated modular semantic static analysis to attempt to prove that the program under analysis does not misbehave and conforms to its specification (lightweight functional correctness).

% ======================================================================
\section{Blending Formal And Agile Development}
\label{sec:blending_formal_and_agile_development}

% principles of agile methodologies
AMs are all based on a highly iterative and incremental process.
They share a common approach on team management, customer relation, simplification and removal of unnecessary artifacts and activities, they rate working solutions and customers satisfaction as the most important indicators considered during development.
The Agile Manifesto~\cite{Beck2001} summarizes the philosophy and principles shared by all AMs.

% verification practices in agile methodologies
The Agile Manifesto does not suggest any specific development techniques, it describes AMs from a very high abstraction level.
Nevertheless, most AMs share a similar approach to artifact verification.
Functional test suites, used as precise requirements or user stories definition, are applied in DSDM~\cite{Stapleton1997}, XP and Crystal Clear~\cite{Cockburn2004}. 
Unit tests and TDD are mentioned, and often included, in most of the AMs defined so far.

% abstraction in test suites
High level test suites, composed of functional tests, describe requirements and bind together requirement documentation and system behaviour: an informal requirement is supported by a set of functional tests. 
The associated development cycle is one iteration long. 
Low level test suites, composed of unit tests, describe the module behaviour and bind together code documentation, code and the contract imposed on the module: unit tests represent the contract and the documentation. 
The associated development cycle is less than one day (as short as a few minutes).

% refactoring in iterative development
Continuous refactoring~\cite{Fowler1999} enables iterative and incremental development, since the system architecture and design, whether explicitly specified or not, must be cleaned up and modified, to accommodate new requirements, services and modules.
Refactoring is possible and feasible because of the safety nets provided by the test suites, used as regression test suites.

\subsection{\mole Practices}

% objectives, repeated
Our objective is to formally specify only parts of the system under development, in an incremental iterative process.
Formal artifacts must coexist with informal ones: requirements, expressed as functional tests or as formal specifications, handmade and generated tests, documentation, source code.
The development process needs to guarantee consistency over all the involved artifacts.

% composition of the two worlds
Both Test Driven Development (TDD) and classical Formal specification Driven Development (FDD) guarantee consistency on all the involved artifacts. 
Both TDD and FDD support iterative development.
The development cycle adopted in TDD is shown in Figure \ref{fig:tdd_diagram} and the one adopted in classic FDD is shown in Figure \ref{fig:fdd_diagram}.
Unit tests used in TDD are substituted  in FDD by formal specifications supported by the verification environment; in our case the verification environment is the generated unit test suite.
Formal specifications replace unit tests: they specify, document and verify the system under development.

\begin{figure}[ht]
  \begin{minipage}[b]{0.48\linewidth}
    \centering
    \includegraphics[scale=0.5]{diagrams/TDD_classic}
    \caption{TDD development cycle.}
    \label{fig:tdd_diagram}
  \end{minipage}
  \hspace{0.02\linewidth}
  \begin{minipage}[b]{0.48\linewidth}
    \centering
    \includegraphics[scale=0.5]{diagrams/FDD_classic}
    \caption{FDD classic development cycle.}
    \label{fig:fdd_diagram}
  \end{minipage}
\end{figure}

% composition is not enough
FDD is used for the formal specified parts of the system, whether TDD is used for the remaining ones: a ``composition'' of formal and agile verification practices.
Yet there are several problems that still need to be addressed.
No communication between the two worlds is permitted, it is not defined how to formally specify a module after it has been implemented and verified in a non formal way.
A module only partially specified is a sort of hybrid that must be treated accordingly.
Finally, in the FDD cycle shown in Figure \ref{fig:fdd_diagram}, the implicit assumption is that it is always possible to identify the correct formal specifications of the system, but this is not always straighforward.
Identifying the correct formal specification is a difficult task, it has to be supported and verified by the development process.

% FDD with handmade tests
A partially specified module cannot be verified properly only through automatically generated tests.
Handmade tests have to be used together with generated ones.
Handmade tests verify complex behaviours that are informally specified, or complex behaviours that are formally specified but cannot be replicated by the formal verification tools, because of the limits of the verification tools themselves. 
The constraints imposed by formal verification tools~\cite{Glass2002} must be taken into account when deciding whether supporting the code with handmade tests, and verification tools based on automatically generated tests are not an exception.
The development cycle able to blend formal specifications, handmade tests and code development is shown in Figure \ref{fig:fdd_hand_made_tests_diagram}: the formal specification drives the handmade tests. 
Both handmade tests and the formal specification drive the code development.

\begin{figure}[ht]
  \begin{minipage}[b]{0.48\linewidth}
    \centering
    \includegraphics[scale=0.5]{diagrams/FDD_hand_made_tests}
    \caption{FDD development cycle with handmade tests.}
    \label{fig:fdd_hand_made_tests_diagram}
  \end{minipage}
  \hspace{0.02\linewidth}
  \begin{minipage}[b]{0.48\linewidth}
    \centering
    \includegraphics[scale=0.5]{diagrams/TDF}
    \caption{TDF development cycle.}
    \label{fig:tdf_diagram}
  \end{minipage}
\end{figure}

% test driven formal specification
The opposite path still needs to be covered, moving incrementally from a tested code to a formally specified code.
In this case, new handmade tests and existing working code drives the formal specification, this is what we call Test Driven Formal specification (TDF).
The resulting development cycle code is shown in Figure \ref{fig:tdf_diagram}, TDF is used to incrementally add formal specifications to an existing working system.

% all together blends into a mole
When the \mole verification practices are used together the initial objectives are achieved: an incremental and iterative process, where development pace is defined by very short development cycles similar to TDD; the complete freedom to decide what is formally specified; the development cycles and the associated verification practices maintain the corresponding artifacts involved consistent, the correct development cycle is determined by the formality of the artifacts involved, ranging from pure development cycles, shown in Figure \ref{fig:tdd_diagram} and \ref{fig:fdd_diagram}, to blended ones, shown in Figure \ref{fig:fdd_hand_made_tests_diagram} and \ref{fig:tdf_diagram}.

% =====================================================================
\section{\mole Verification Practices Applied}
\label{sec:a_complex_case_study}

\subsection{Data Stream, Protocol Description and Protocol Specification}
\label{subsec:data_stream_and_protocol_description}

% protocol layers
The \SB protocol is asynchronous and packet-based; the packets have a fixed length.  
The protocol is separated in multiple layers to ease its implementation in the driver.
The layers identified are the following: (1) packet byte structure: the internal representation of the packet; (2) packet info structure: the meaningful fields contained in the packets; (3) single packet rules: the content acceptable values, and how they influence each other; (4) packet sequence rules: the content acceptable values, based on values of previously received packets; (5) packet input output asynchronous rules: the content acceptable values and reaction constraints on output packets, based on previously transmitted input packets.
The first three layers are analyzed in the case study, considering only the output sensor data packet.

% packet structure
The packet byte and info structure of the output sensor data packet reflect the board capabilities and built in sensors. 
A single packet has to accommodate various types of data: fast, medium and slow data rate streams, together with metadata describing the sensors and the board state.
The internal structure of the packet is strictly fixed.

% packet internal structure: collection of frames
A packet is internally composed by 82 frames. 
A frame accommodates data from the fast and medium data rate streams (4 fast data rate stream and 8 medium data rate stream channels) and theirs associated metadata.
%Fast, medium and slow data rates have not a common multiple; a packet, capable to accommodate outputs at different rates, in a fixed structure, relies in two forms of redundancies, for slow and medium data rates:
%\begin{itemize}
%\item Slow data rates: slow data is located at the beginning of each packet, the data is simply repeated (it is equal to the previous packet in the stream) when no new data is available.
%\item Medium data rates: the 8 medium rate channels are multiplexed together into the medium data rate data sample of each frame. The medium data rate sample channel is a piece of info extracted from the frame description (frame metadata), including the possibility to have no data at all. 
%E.g. a frame contain a medium data rate sample from channel 0, the following frame a sample from channel 5, the following frame no sample at all, etc.
%\end{itemize}

% single packet rules
The single packet rules delimit the boundaries of the values obtainable from the packet: each defined sensor represented in the packet, as well as the metadata describing the \STSB and the packet and frame contents, are constrained by a range of acceptable values.
There are also rules affecting more than one value, and rules specifying a correct sequence of frames.

%\subsection{Data stream and protocol specification}
%\label{subsec:data_stream_and_protocol_specification}

% protocol specification
The specifications of the protocol are distilled and refined incrementally.
The protocol is divided in various (thin) layers, and each of the layer is verified with a different approach: some of the layers are specified formally.

\paragraph*{Packet byte structure specification}

% verification
Verification is obtained through an handmade test suite, composed of unit and integration tests; the tests specifies the behaviours of the implementation, which is capable of recognizing the proper packet structure in a binary data stream. 
%The implementation is tricky: it searches for repeated patterns into a byte stream to be sufficiently confident on how the stream is decomposed into packets.
The packet byte structure is not stable: the internal byte structure changed several times during development.
Therefore, packet byte structure is informally specified.

% development cycles
The test suites are used to verify board simulators (low-level), but they cannot be used to check properties at runtime.
Code and tests are implemented using TDD.

\paragraph*{Packet info structure specification}

% verification
Verification is provided by handmade test suite, JML annotations and generated unit test suite. 
It has been built starting from handmade tests only (TDD development cycles).
Later on JML annotations have been added incrementally, using a mix of \mole practices: to move from handmade tests to formal specifications, TDF have been initially used; than all the \mole practices have been used according to the artifacts involved.

% development cycles
The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check the formally specified properties at runtime.
Code and tests have been initially implemented using TDD, than using all the \mole verification practices.

\paragraph*{Single packet rules}

% verification
The verification is provided exclusively by the formal specification with JML language, combined with the generated unit test suite and handmade test suite to verify the most complex behaviour.
It is built starting from JML annotations, supported by handmade tests when needed.

% development cycles
The test suites are used to verify board simulators (high-level), RAC and JML annotations are used to check most of the rules at runtime.
Code and tests are implemented using both forms of FDD.

\paragraph*{On simulators}
%\label{subsec:on_simulators}

% why simulators are useful
Simulators are used to test parts of the system minimizing dependencies: a simulator can be used by upper layers, with no need to provide the functionalities of the lower layers. 
%The simulators proved essential during the development, since the physical board was not available.

% simulators description
The layer structure is not matched by the corresponding simulators.
The driver API is splitted in two abstraction layers: the general high-level interface, that exposes the main functionalities of the board and the main contents of the packets, and the lower level implementation that parses the data streams in and out the board, meant to translate the higher level instructions and data in properly formed packets.
The simulators are built according to these abstraction layers.
The high-level simulator implements the interfaces providing the methods to deal with an abstracted \SB.
The low-level simulator is capable to rebuild the \SB data streams: the in and out data streams are built exactly as the sensor board is expected to parse or generate.

\subsection{JML Specification Examples}
\label{subsec:a_jml_specification_example}

% a simple specification example
The JML specifications are of varying complexities.  
Some of them are rather simple, focusing on constraints that should hold when calling a method (the preconditions) and constraints on the return value (the very basic form of postconditions). 
In listing \ref{lst:jml_example_simple} a simple JML specification example is shown, the method \lil{getTemperature} is declared \lil{/*@ pure @*/}, which means that it cannot change the state of any object (a postcondition); the specification also constraints the return value with a lower and upper bound (another postcondition).

\begin{lstlisting}[
  caption={A simple specification with JML annotation: simple postconditions.},
  label=lst:jml_example_simple,
  float=htbp]
  /*@ 
      ensures \result >= -880; 
      ensures \result <= 2047; 
  @*/ 
  /*@ pure @*/ short getTemperature();
\end{lstlisting}

\sloppy

% a complex specification example
The complex specifications usually focus on properties regarding the behaviour of a whole object.
In listing \ref{lst:jml_example_invariant} a complex invariant example is shown; an invariant is a property maintained during the life cycle of an object, more precisely, an invariant is assumed on entry and guaranteed on exit of each method of an object. 
The invariant is constraining the number of samples for each medium data rate streams: the total number of streams is \lil{Frame.ADC_CHANNELS}, the total number of frames is \lil{FRAMES}, the constant constraining the number of samples is \lil{FRAMES/Frame.ADC_CHANNELS+1}, meaning that the samples contained in a frame are fairly distributed on the channels.
The valid samples are counted parsing all the frames contained in a packet, selecting only the matching valid samples.  
A medium data rate stream sample is considered valid when \lil{isADCActive()} method returns \lil{true}.

\fussy

\begin{lstlisting}[
  caption={A specification with JML annotation: a complex object invariant which constraints 
    the number of samples for each medium data rate stream channel.},
  label=lst:jml_example_invariant,
  float=htbp]
  /*@ 
      invariant ( 
        \forall int channel; 
            0 <= channel && 
            channel < Frame.ADC_CHANNELS; ( 
          \num_of int i; 
              0 <= i && 
              i < (FRAMES-1); (
            (getFrame(i).isADCActive()) && 
            (getFrame(i).getADCChannel() == channel)
          ) 
        ) <= (FRAMES / Frame.ADC_CHANNELS + 1) 
      ); 
  @*/
\end{lstlisting}

\subsection{Test Cases}
\label{subsec:test_cases}

% unit tests kinds
The unit test cases that verify the protocol driver are of two kinds: handmade unit tests and automatically generated unit tests based on JML specifications; they are complementary and built to be used together.

% unit tests effectiveness
The test effectiveness is evaluated for each test suite; the evaluation is carried out in section \ref{subsec:test_cases_retrospectives}.
The test effectiveness evaluation considers three elements: effective results on piloting the real board (quantitative), code coverage
(quantitative), development help and usefulness (qualitative).

\paragraph*{Handmade tests}

% the driver package structure
The package structure of the driver is shown in Figure \ref{fig:class_diagram_main}: two independent packages are defined (\lil{Stream} and \lil{Driver}).  
The packages are abstract, that is, they mainly contain abstractions; in Java language this is translated into a package which contains mainly abstract classes or interfaces.  
The dependency between the two abstract packages is not direct, it is realized through an implementation (\lil{StreamDriver}).
This is needed to maintain a high decoupling of the packages, and is the result of applying the dependency inversion principle~\cite{Martin1996}.

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__Main}
  \caption{Main packages class diagram.}
  \label{fig:class_diagram_main}
\end{figure}

% the test package structure
The test package structure, shown in Figure \ref{fig:class_diagram_test}, reflects and mimics the code structure; the tests for an abstract package are abstract, and implemented by the package that tests a corresponding system implementation. 
This is a well known test pattern (the Abstract Test Pattern~\cite{Thomas2004}) used to test that the contracts defined in the abstractions are respected in all the implementations.  
For instance, package \lil{DriverT} contains abstract tests for the abstractions of package \lil{Driver}, package \lil{StreamDriverT} inherits the abstractions of \lil{DriverT} and makes them concrete, to test the corresponding implementation \lil{StreamDriver}; package \lil{StreamDriverT} also contains stand alone tests written specifically for the implementation \lil{StreamDriver}.

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__Test}
  \caption{Test packages class diagram.}
  \label{fig:class_diagram_test}
\end{figure}

\paragraph*{Generated tests}

% the generated test package structure
The JML specifications are used to generate tests.
The resulting test package structure is closely related and reflects the overall package structure; each package have a corresponding generated test package.

%\begin{figure}[htb!]
%  \centering
%  \includegraphics[scale=0.7]{UML_model/Class_Diagram__Structure__GeneratedTests}
%  \caption{Generated test packages class diagram.}
%  \label{fig:class_diagram_generatedtest}
%\end{figure}

\paragraph*{On simulators}

% simulator roles in tests
Simulators are adopted in the handmade test suits as stubs.
The resulting structure is shown in Figure \ref{fig:class_diagram_streamdriver_test}.  
For instance, in Figure \ref{fig:class_diagram_streamdriver_test} the test suite \lil{StreamDriverT} is using the \lil{SimulatorStream} to properly simulate the \lil{Stream} parsed by \lil{StreamDriver}, which is the system under test.

\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.4]{UML_model/Class_Diagram__Structure__StreamDriverT}
  \caption{StreamDriver test packages class diagram: use of the stream
    simulator.}
  \label{fig:class_diagram_streamdriver_test}
\end{figure}

% simulator adoption
The simulators are adopted in both the handmade and the generated test suite; both the suites need a proper set of stubs in order to be executed.

\subsection{Test Cases on the Job}
\label{subsec:test_cases_on_the_job}

% test suites sizes
The handmade test suite is composed by over 140 tests, while the generated one is 100 times bigger, totalling over 14000 tests.
The large number of generated tests are explained looking closely at how the generated tests are created by the JML framework. 
The generated tests explore the possible input combinations on public methods, combining the type data ranges that are specified for the test suite (see \cite{Cheon-Leavens02} for more details on how the JML framework works on generating unit tests). 
For instance, let's suppose a method \lil{m(int p1, int p2)} must be tested in a class \lil{C}; the data ranges specified for type \lil{int} are \lil{\{1,2,3\}} and for type \lil{C} are the instances \lil{\{c1,c2,c3,c4\}}. 
The JML framework generates $24 = 3 \times 3 \times 4$ tests, since they are the combination of the data ranges involved (the data range of the parameters and the data range of the type owning the method under test). 
Nevertheless, not all the generated tests are executed: a generated test is not executed, unless the preconditions of the called method are satisfied. 
Thus, depending on the preconditions specified for a method, the number of the generated tests that are actually executed in the test suite are significantly smaller compared to the total number of generated tests.

% test execution time, values
Test suites execution time are different: the handmade test suite, with the runtime assertion checking active, runs completely in less than one minute on an average PC, while the generated test suite runs in more than 20 minutes (the rough ratio is $ 1 : 60 $). 
Most of the time is spent on setting up the suite; only considering the real execution of the tests, with no suite setup, the execution time reduces to less than 4 minutes (the rough ratio is $ 1 : 10 $).

% test effectiveness
Code coverage metrics (white-box testing) are used to compare in a quantitative way the effectiveness of the test suites. 
This is only an indicator, since it is accepted that code coverage alone cannot asses the quality of a test suite\cite{Marick1999,Chockler2006}.
Various coverage metrics exist, the test suites are analyzed with statement code coverage (one of the simplest types)\footnote{The tool used to obtain statement code coverage metrics is \myhref{http://emma.sourceforge.net/}{Emma}.}: statement coverage reports whether each statement is encountered during execution. 
The coverage results are reported in the Table \ref{tab:statement_code_coverage}.

\begin{table}[htbp]
  \caption{Statement code coverage result, categorized by source package.}
  \label{tab:statement_code_coverage}
  \begin{center}
    \begin{tabular}{|c|c|c|c|}\hline
      & \textbf{\textit{hand}} & \textbf{\textit{generated}} &
      \textbf{\textit{total}} \\\hline
      \textbf{\textit{Driver}} & $15.9 \%$ & $6.3 \%$ & $15.9 \%$ \\\hline
      \textbf{\textit{StreamDriver}} & $76 \%$ & $79.7 \%$ & $88.5 \%$ \\\hline
      \textbf{\textit{SimulatorDriver}} & $64.9 \%$ & $44.2 \%$ &
      $71.2 \%$ \\\hline
      \textbf{\textit{Utility}} & $98.1 \%$ & $74.5 \%$ & $98.1 \%$ \\\hline
    \end{tabular}
  \end{center}
\end{table}

%\subsection{Retrospective on test cases effectiveness}
%\label{subsec:test_cases_retrospectives}

% <100 coverage measures, utility methods
The coverage measures reported in Table \ref{tab:statement_code_coverage} do not reach 100\%. 
This is because of the effects of both non public utility methods, and abstract methods in interfaces.
The effect of non public methods (package Java visibility) is that these methods are taken into account as public and protected ones are, during code coverage calculation; these methods are not part of the interface, the system does not depend on them, they are only used in object initialization or in object setups performed during unit tests. 
%In fact there are package and private code blocks that are not reachable by construction from public and protected methods, and this is a precise characteristic reflecting the code architecture, meant to ease class instances setup and configuration for testing purposes (white box testing). 
%The used statement coverage tool is not able to distinguish these code blocks, so this effect is unavoidable. 
The effect is mainly seen in low \lil{SimulatorDriver} code coverage figures: the \lil{SimulatorDriver} package has many utility non public methods.

% <100 coverage measures, interfaces
Regarding interfaces, an interface contains no statements, so when an interface method is called, it is the method of the implementation class used that is actually covered.
This effect is visible in \lil{Driver} coverage figures: \lil{Driver} package contains interfaces (that are not counted at all) and some very simple real class (exceptions) that are not thoroughly tested. 
%The net result is extremely low statement code coverage figures because the very simple classes are the only ones that are actually covered.

% code coverage interpretation
The statement code coverage figures show that the test suites do a good job on the most important package, the \lil{StreamDriver}. 
The generated test suite reaches almost $ 80 \% $, while the handmade tests have only a slightly lower coverage ratio: $ 76 \% $. 
The combination of test suites raise the coverage ratio to $ 88.5 \% $. 
This is a clear indication that one test suite is not completely overlapping the other: the intuition that test suites are not alternatives, but have to be used together to achieve the higher benefits, is confirmed.
A similar indication is provided by \lil{SimulatorDriver} package as well.

% handmade unit test vs generated unit test
The qualitative difference of the two test suites are understood observing how the suites are built. 
An handmade unit test usually requires objects initialization, one or more method calls to the modules under test, and the following assertions to verify the expected state change on the modules involved. 
A generated unit test has a more focused and limited scope: only one method call is performed in each test, therefore it is difficult to explore complex behaviours. 
%Theoretically with a complete specification, and specifying large enough data ranges, the generated tests achieve the same expressive power as the handmade ones; but the specification is not complete, and a large enough data ranges is long to develop.
The test suites give their best when used in a combined approach: the generated test suite, checking simpler and formalized behaviours, and the handmade, checking the more complex ones, whether formalized or not.

\paragraph*{Effectiveness on board delivery}

% tests and the first prototype, board side
The first \STSB prototype builds packets with no data, but the packets are built in a way not consistent with the specification.  
The board errors are detected by using the driver implementation, the \lil{StreamDriver} package.

% tests and the second prototype, driver side
The second prototype builds packets with values taken from real sensors installed on the board.  
The only sensors that are not working are the sensors devoted to the fast rate data streams (audio sensors).  
One error needed to be corrected in the driver implementation, because of a (rare) combinations of conditions not initially covered by the test cases, but that occasionally showed up during real use.
%One error was discovered in the low-level system driver used by the developed driver (the system driver is outside our development scope): to eliminate the error, an upgrade of the low-level driver to a new beta version was necessary. 
%No other errors were detected on the driver side.

% tests and the second prototype, board side
The second prototype respected the packet byte structure specification, but was not fully compliant to the packet info structure specification and the single packet rules.
It also randomly generated completely invalid packets when overheated.
The formal JML specifications and the runtime assertion checker correctly identified these errors.
4 protocol errors regarding the packet info structure specification, and 2 protocol errors regarding the single packet rules were found, as well as the invalid packets.

\section{Retrospective on \mole Practices Effectiveness}
\label{subsec:test_cases_retrospectives}

% preconditions on projects
We believe that \mole practices are relatively easy to introduce and
use.  Many kinds of projects could benefit from these practices, not
only those that have unstable requirements but also demands for a
certain level of formality.  Even projects with no requirements for
formal specification at all benefit from these practices, since some
simple or obvious system properties are easy to specify and, though
their introduction, there is a corresponding reduction in the test
suite development effort.

% precondition on formal methodology and associated formal suite
The cornerstone of the practices is the formal methodology and its
supporting tools, and the precondition on the tools' utility is not
always easily fulfilled.  Methodologies that require a complete
formalization are nearly impossibile to apply.  In addition, it is not
clear if a tool suite that does not provide a runtime verification,
either via runtime checking or simulation, can be broadly used
succesfully, nor it is clear whether formal methods that are not based
on the contract paradigm are useful in this context.  These are open
questions that will need to be addressed in the future to determine
the broad, concrete applicability of the \mole practices.

% when to use what
There are no strict rules to determine which of the \mole practices to
use when facing a specific problem.  We have found that experience
trumps cookbooks, so we can only give some advice.  We suggesting
using TDD there is no need for any specifications at all, FDD with
generated tests if the specification are not too difficult and system
scenarios is simple enough to be covered by generated tests, FDD with
hand made tests if the scenarios are complex, and TDF when the
specification is complex.  We feel that the experience we have
gathered is not yet enough to distill richer idioms or patterns---such
is future work.

% what is working and what is not
Upon reflect, we find that the end result obtained via combining FDD
with hand made tests is remarkable, as it is as natural to work with
as TDD.  FDD with generated tests is sometimes annoying because of the
time needed to run the full test suite, thus often only part of the
test suite is executed so that the development loops do not take too
long.  TDF is great for complex preconditions, but really difficult
with postconditions, especially if the specification is written for
existing code.  When writing after code, the corresponding test is not
expected to fail, hence the feedback is limited and the contract is
often not as strong as needed.  In this specific case, TDF
effectiveness is limited.

\section{Conclusions}
\label{sec:conclusions}

% the problem
In this paper we focus on a specific set of development challenges
with which neither AMs nor FMs are completely confortable.  The
challenges are characterized by unstable requirements combined with
artifacts that must be formally specified and verified all while the
development team is constrained by deadlines that cannot be postponed.
We propose a blend of agile and formal engineering practices, enclosed
by an iterative and incremental development process: the \mole
development process.

% Mole
The verification practices described in this paper recreate the fast
feedback development environment we find in Test Driven Development in
the presence of formal and informal artifacts.  A total of four
practices are presented---two of them are conservative, following
closely the development practices of formal methodologies and agile
methodologies, while the other pair are innovative, blending together
elements from both worlds.

% Mole at work
The practices have been applied succesfully to develop a driver for a
custom embedded sensor board equipped with board simulators and
protocol verifiers.  And, while the protocol and the board
specifications were informal and unstable, a working software product
was needed as soon as possible, with incrementally added
functionality, and ensuring that the software and hardware products
were linked via the last specifications of the protocol and the board.
\mole verification practices enabled us to keep in sync formal
specifications (JML annotations), informal specifications (test
suites), and source code through very rapid, short development cycles.
We developed handmade test suites and generated test suites and JML
annotations, all of which successfully supported the verification of
the software driver and the simulators, as well as the hardware board
when it was delivered.  In fact, the resulting system had no flaws, as
it worked correctly the very first time it was plugged in to the
prototype hardware.

% extending Mole
The practices we describe are general and are applicable to different
development languages and formal methodologies.  One important
constraint is on the formal methodology is that it must cope with
partial specifications.  Another constraint is on the tools available,
which must verify system properties at runtime and partially automate
the verification process of the formalized properties.

TODO POP!

%\bibliographystyle{lni}

\bibliography{extra,abbrev,ads,category,complexity,hypertext,icsr,knowledge,languages,linguistics,meta,metrics,misc,modeling,modeltheory,reuse,rewriting,softeng,specification,ssr,technology,theory,web,upcoming,upcoming_conferences,conferences,workshops,verification,escjava,jml,nijmegen}

\end{document}

%=====================================================================
% hi Joe

% what about having the chat next week?
% on thursday, maybe?

% here are my comments on the reviewers:

% ---------------------------- REVIEW 1 --------------------------
% OVERALL RATING: 3 (strong accept)

% > This paper successfully bridges the gab between both worlds. I would be i=
% nteresting to read at bit more about the gained experience with the techniq=
% ues based on the presented case study (what was easy what difficult).

% agree. to be included

% > Furthermore I recommend that the authors work out a bit more detailed the=
%  prerequisites to apply a mixture of formal and agile techniques (e.g. by m=
% eans of a list of "thesis"). This would stimulate the discussion!

% agree. to be included. it's already there, but not clear and polished enoug=
% h

% ---------------------------- REVIEW 2 --------------------------
% OVERALL RATING: 2 (accept)

% > The authors use the (somewhat formal) JML technology and introduce their =
% TDF-methodology as a variant of more classic development cycles.

% why "somewhat"?

% > Content-wise it must be said that, given the amount of material in the te=
% xt, the conclusions drawn and lessons learnt are a bit 'thin'. E.g. Does TD=
% F really work? what is your experience?

% partial agree. to be expanded, in the line of the previous comments

% > It seems that the format of the printed area (text height and width) is n=
% ot in accordance with the other conference contributions. When reformatted,=
%  the already long paper will probably get longer. The (too large) wordiness=
%  of the paper could perhaps be reduced somewhat to compensate.

% uh
% I used thir template... I'm "almost" sure
% I'll double check this
% maybe some following format options

% > page 1, abstract: 'not always see eye to eye' seems an inappropriate way =
% of stating 'do not meet face to face'.
% > page 1, abstract end: highly-spices -> highly-spiced
% > page 2, line 2: citation problem
% > page 2, line 12: 'ther'?
% > page 3, line 8 and 10: appliable in -> applicable to?
% > page 3, line 8: yeilding -> yielding
% > page 3, middle: .... the handmade test suites are automatically generated=
%  .... ?
% > page 3, the same section: the last two sentences seem crippled; split the=
% m up!
% > page 6, line -5: Identify -> Identifying
% > page 9, line 1: confidant -> confident
% > page 15, line 6: inoovative -> innovative
% > page 15, line 12: in sink -> in sync
% > page 17, last line: litt ref in disorder

% ok

% > Summing up the editorial part: Please use a spelling checker and perhaps =
% some review by a native speaker
% could help sharpening (and shortening) some sentences.

% well... I've used a spell checker!
% but we have to reread it once the modifications are in place

% ---------------------------- REVIEW 3 --------------------------
% OVERALL RATING: 2 (accept)

% > they do not mention Liu's work on the SOFL foraml engineering method for =
% software development that have been xtensively studies over the last 15 yea=
% rs with the aim to solve the similar problem. In particular, the recent pap=
% er on the application of SOFL to achieve agile process has provided a very =
% different approach from the mentioned approaches
% > Shaoying Liu, ``An Approach to Applying SOFL for Agile Process and Its Ap=
% plication in Developing a Test Support Tool'', Innovations in Systems and S=
% oftware Engineering, Springer London, 22 December, DOI 10.1007/s11334-0114-=
% 3.

% ok
% to be read and analyzed

% ---------------------------- REVIEW 4 --------------------------
% OVERALL RATING: 1 (weak accept)

% > I did find it disturbing that the paper never clearly distinguishes betwe=
% en structural and functional testing . This is an
% important distinction as it potentially evidences different types of proble=
% ms.

% I'll rewrite some parts to make the distinction clear

% > Moreover, there is a distinct difference between automatically generated =
% tests and manually produced tests as to how well they apply to one or the o=
% ther.

% the distinction is clear in the paper... I don't understand this

% > The overall description of how AM and FM are merged is not that persuasiv=
% e, as a result. In particular, Figure 2 does not
% convey much nor profound value in its current incarnation and description.

% mmhh
% not much that can be done here... I think
% how can we be more "persuasive"?

% > It would help if the authors would more specifically address what type of=
%  tests the JML specification helps generate. In the
% case study, it is a structural coverage type test.

% mmhh
% I've explained this
% it would take me pages to go into specific details... and I've cited
% works where this is explained

% > All that said, it is a bit unclear as to where the presented work is actu=
% ally pushing the boundaries of the current state of
% the field, other than in actually presenting the approach as an FM+AM hybri=
% d.

% we are actually talking about FM+AM hybrids... that IS the filed,
% so... what's this question about?

% > As a suggestion, then, it could be good if the verbosity and obtusity of =
% the article is reduced

% ok...

% > and a more formal and incisive treatise is presented.

% no idea how to do this
% nor I'm sure I really want to do this

% --=20
% Vieri del Bianco